\input ../pree.tex

\begin{document}

\title{Abelian categories - some results}
\date{}
\maketitle

\section{Directed classes and filtered categories}
\noindent
In this section we introduce the notion of directed class and then we generalize it to the categorical setting.

\begin{definition}
Let $I$ be a class equipped with a reflexive and transitive relation $\leq$. We say that $I$ is \textit{a directed class} if for any $i$, $j\in I$ there exists $k\in I$ such that $i\leq k$ and $j\leq k$.
\end{definition}
\noindent
Let $I$ be a class equipped with a reflexive and transitive relation $\leq$. It is standard {\cite[page 11]{Maclane}} to view $I$ as a category with at most one arrow between any two objects.

\begin{definition}
Let $I$ be a category. Suppose that the following conditions are satisfied.
\begin{enumerate}[label=\textbf{(\arabic*)}, leftmargin=1.5em]
\item For any objects $i$, $j\in I$ there exists an object $k\in I$ and a diagram
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=2em, column sep=2em,text height=1.5ex, text depth=0.25ex] 
{     &   k &                    \\
      i &  &  j                       \\} ;
\path[->,font=\scriptsize]  
(m-2-1) edge node[auto] {$ $} (m-1-2)
(m-2-3) edge node[auto] {$ $} (m-1-2);
\end{tikzpicture}
\end{center}
\item For any pair of parallel morphisms in $I$
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{  i &    j                       \\} ;
\path[->,font=\scriptsize]  
(m-1-1) edge[transform canvas={yshift=0.5ex}] node[auto] {$ $} (m-1-2)
(m-1-1) edge[transform canvas={yshift=-0.5ex}]  node[auto] {$ $} (m-1-2);
\end{tikzpicture}
\end{center}
there exist an object $k\in I$ and a morphism $j\ra k$ such that, the diagram 
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{  i &    j                   & k     \\} ;
\path[->,font=\scriptsize]  
(m-1-1) edge[transform canvas={yshift=0.5ex}] node[auto] {$ $} (m-1-2)
(m-1-1) edge[transform canvas={yshift=-0.5ex}]  node[auto] {$ $} (m-1-2)
(m-1-2) edge  node[auto] {$ $} (m-1-3);
\end{tikzpicture}
\end{center}
is commutative.
\end{enumerate}
Then we say that $I$ is \textit{a filtered category}.
\end{definition}

\begin{fact}\label{fact:filteredisgeneralizationofdirected}
Let $I$ be a class equipped with a reflexive and transitive relation. Then $I$ is a directed class if and only if $I$ viewed as a category is filtered.
\end{fact}
\begin{proof}
Left to the reader.
\end{proof}

\begin{fact}\label{fact:imageoffilteredinclass}
Let $I$ be a filtered category and $J$ be a class equipped with a reflexive and transitive relation $\leq$. Suppose that $F:I\ra J$ is a functor. Then the image of $I$ under $F$ is a directed subclass of $J$.
\end{fact}
\begin{proof}
Left to the reader.
\end{proof}

\begin{definition}
Let $\cC$ be a category and $I$ be a filtered category (directed class). A colimit of an $I$-indexed some diagram $I\ra \cC$ is called \textit{a filtered (directed) colimit}.
\end{definition}

\section{Subobjects and objects of finite type}

\begin{definition}
Let $\cC$ be a category and $X$ be an object in $\cC$. Monomorphisms $X_1\hookrightarrow X$ and $X_2\hookrightarrow X$ are \textit{equivalent} if there exists a commutative triangle
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=2em, column sep=1em,text height=1.5ex, text depth=0.25ex] 
{     X_1     &   &  X_2        \\
              & X &  \\} ;
\path[->,line width=1.0pt,font=\scriptsize] 
(m-1-1) edge node[above] {$\cong $} (m-1-3);
\path[right hook->,line width=1.0pt,font=\scriptsize] 
(m-1-1) edge node[below] {$ $} (m-2-2)
(m-1-3) edge node[left] {$ $} (m-2-2);
\end{tikzpicture}
\end{center}
in which horizontal arrow is an isomorphism. The collection $\mathrm{Sub}(X)$ of equivalence classes of monomorphisms having $X$ as a target is called \textit{the class of subobjects of $X$}.
\end{definition}
\noindent
Let $X$ be an object of a category $\cC$ and $X'\hookrightarrow X$ be a monomorphism. By abuse of notation we say that $X'$ is an subobject of $X$ and by this we understand the subobject of $X$ represented by a monomorphism $X'\hookrightarrow X$. We also write $X'\subseteq X$. Next suppose that $X_1\subseteq X$ and $X_2\subseteq X$ are subobjects of $X$. We write $X_1\subseteq X_2$ if there exists a commutative triangle
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=2em, column sep=1em,text height=1.5ex, text depth=0.25ex] 
{     X_1     &  &  X_2        \\
              & X&  \\} ;
\path[right hook->,line width=1.0pt,font=\scriptsize]
(m-1-1) edge node[above] {$ $} (m-1-3) 
(m-1-1) edge node[below] {$ $} (m-2-2)
(m-1-3) edge node[left] {$ $} (m-2-2);
\end{tikzpicture}
\end{center}
This defines a partial order on the class $\mathrm{Sub}(X)$.\\
Now we investigate important notion of well-powered categories. For this we introduce this it with a company of several other significant concepts. 

\begin{definition}
A category $\cC$ is called \textit{well-powered} if $\mathrm{Sub}(X)$ is a set for every object $X$ in $\cC$.
\end{definition}

\begin{definition}
Let $\cC$ be a category such that a morphism in $\cC$ that is simultaneously a monomorphism and an epimorphism is an isomorphism. Then we say that $\cC$ is \textit{balanced}.
\end{definition}

\begin{definition}
Let $\cC$ be a category. A class $\cG$ of objects of $\cC$ is called \textit{a class of generators for $\cC$} if for any pair of distinct and parallel arrows 
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=2em,text height=1.5ex, text depth=0.25ex] 
{ X&  Y  \\} ;
\path[->,line width=1.0pt,font=\scriptsize]
(m-1-1) edge[transform canvas={yshift=0.5ex}] node[above] {$ f$} (m-1-2)
(m-1-1) edge[transform canvas={yshift=-0.5ex}] node[below] {$ g$} (m-1-2);
\end{tikzpicture}
\end{center}
there exists $G\in \cG$ and a morphism $h:G\ra X$ such that $f\cdot h\neq g\cdot h$.
\end{definition}

\begin{proposition}\label{proposition:criterionforwellpoweredcategories}
Let $\cC$ be a balanced, locally small category that admits fiber products. Assume that $\cG$ is a \textbf{set} of generators for $\cC$. Then $\cC$ is well-powered. 
\end{proposition}
\begin{proof}
Fix an object $X$ of $\cC$. Then every subobject $X'$ of $X$ gives rise to a set
$$\mathrm{Factor}(X') = \big\{f\in \Mor(\cC)\,\big|\,\mathrm{dom}(f)\in \cG,\,\mathrm{cod}(f) = X\mbox{ and }f\mbox{ factors through }X'\big\}\subseteq \prod_{G\in \cG}\Mor_{\cC}(G,X)$$
It suffices to prove that 
$$\mathrm{Sub}(X)\ni X'\mapsto \mathrm{Factor}(X')\in \cP\left(\prod_{G\in \cG}\Mor_{\cC}(G,X)\right)$$
is injective (because a class bijective with a set is a set itself). For this assume that $X_1$ and $X_2$ are subobjects of $X$ such that $\mathrm{Factor}(X_1) = \mathrm{Factor}(X_2)$. Since $\cC$ admits fiber products, we derive that there exists a fiber product of $X_1\hookrightarrow X$ and $X_2\hookrightarrow X$. We denote it by $X_1\cap X_2$ and consider it as a subobject of $X$ via the canonical map $X_1\cap X_2 \hookrightarrow X$. By universal property of fiber product we deduce that $\mathrm{Factor}(X_1) = \mathrm{Factor}(X_1\cap X_2) = \mathrm{Factor}(X_2)$. This implies that for every object $Y$ of $\cC$ maps
$$\Mor_{\cC}(X_1,Y)\ra \Mor_{\cC}(X_1\cap X_2,Y),\,\Mor_{\cC}(X_2,Y)\ra \Mor_{\cC}(X_1\cap X_2,Y)$$
induced by $X_1\cap X_2\hookrightarrow X_1$ and $X_1\cap X_2\hookrightarrow X_2$ are injective. Therefore, morphisms $X_1\cap X_2\hookrightarrow X_1$ and $X_1\cap X_2\hookrightarrow X_2$ are epimorphisms. Since they are also monomorphisms and $\cC$ is balanced, they are isomorphisms. Thus $X_1$, $X_1\cap X_2$, $X_2$ represent the same subobject of $X$.
\end{proof}

\begin{definition}
Let $\cC$ be a category and $X$ be an object in $\cC$. \textit{A filtered (directed) family of subobjects of $X$} is a functor $I\ra \mathrm{Sub}(X)$ from a small filtered category (directed set) $I$.
\end{definition}
\noindent
Suppose that $\cC$ is a category, $X$ is an object of $\cC$ and $I$ is a filtered category. Let $I\ra \mathrm{Sub}(X)$ be a filtered family of subobjects of $X$. Then it can be described as a map $I\ni i \mapsto X_i\in \mathrm{Sub}(X)$ such that for every morphism $i\ra j$ in $I$ we have $X_i\subseteq X_j$. For pragmatical reasons we usually use this more explicit description and we view filtered families of subobjects as an indexed families of the form $\{X_i\}_{i\in I}$.

\begin{definition}
Let $\cC$ be a category and $X$ be an object in $\cC$. A filtered family of subobjects $\{X_i\}_{i\in I}$ of $X$ is \textit{complete} if $X = \mathrm{colim}_{i\in I}X_i$
\end{definition}

\begin{definition}
Let $\cC$ be a category and $X$ be an object in $\cC$. Suppose that for every complete filtered family $\{X_i\}_{i\in I}$ of subobjects of $X$ there exists $i_0\in I$ such that $X_{i_0} = X$ for every $i\in I$. Then we say that $X$ is \textit{of finite type}.
\end{definition}

\begin{definition}
Let $\cC$ be a category. We say that $\cC$ is \textit{locally finite} if for every object $X$ there exists a complete filtered family $\{X_i\}_{i \in I}$ of subobjects of $X$ such that $X_i$ is of finite type for every $i\in I$.
\end{definition}

\section{Abelian Categories - definition and statement of the embedding theorem}

\begin{definition}
Let $\cC$ be a locally small category. Suppose that the following conditions hold.
\begin{enumerate}[label=\textbf{(\arabic*)}, leftmargin=1.5em]
\item $\cC$ has a zero object (i.e. an object that is both initial and terminal).
\item $\cC$ has finite limits and colimits.
\item Each epimorphism in $\cC$ is a cokernel of some arrow and each monomorphism in $\cC$ is a kernel of some arrow.
\end{enumerate}
Then $\cC$ is called \textit{an abelian category}.
\end{definition}
\noindent
The definition of an abelian category is taken from \cite{freyd1964abelian}. This differs from the more popular definition {\cite[page 198]{Maclane}} - it does not assume the existence of $\Ab$-enrichment (abelian group structure on sets of morphisms). The two notions are equivalent by {\cite[Theorem 2.39]{freyd1964abelian}}.\\
The next result shows that all elementary properties of categories of modules over a ring hold for general abelian categories.

\begin{theorem}[Freyd-Mitchell embedding]\label{theorem:Freyd_Mitchel_embedding}
Let $\cC$ be a small abelian category. Then there exists a ring $R$ and a full, faithful and exact functor $E:\cC\ra \Mod(R)$.
\end{theorem}
\noindent
This is {\cite[Theorem 7.34]{freyd1964abelian}}.

\section{\textbf{Ab}-conditions in categories}
\noindent
In this section we discuss Grothendieck's $\bd{Ab}$-conditions in categories. The original source of the discussion below is the seminal work \cite{grothendieck1957}. First let us explain that Grothendieck introduces \textit{$\bd{Ab}0$-categories} as additive categories, \textit{$\bd{Ab}1$-categories} as preabelian categories and \textit{$\bd{Ab}2$-categories} as the usual abelian categories. Then he continues with more specific classes of abelian categories and recapitulation of parts of his work is our main task here.

\begin{definition}
A category $\cC$ is \textit{an $\bd{Ab}3$-category} if it is an abelian category and small direct sums in $\cC$ exists.
\end{definition}
\noindent
Let $\cC$ be an $\bd{Ab}3$-category, $\{X_i\}_{i\in I}$ be a family of objects of $\cC$ and let $\big\{f_i:X_i\ra Y\big\}_{i\in I}$ be a family of morphisms in $\cC$. Then we denote by 
$$\sum_{i\in I}f_i:\bigoplus_{i\in I}X_i\ra Y$$
a unique morphism determined by requirement $\left(\sum_{i\in I}f_i\right)\cdot v_i = f_i$ for each $i\in I$, where $v_i:X_i\ra \bigoplus_{i\in I}X_i$ is the canonical inclusion.

\begin{theorem}\label{theorem:descriptionofcolimitsinab3categories}
Let $\cC$ be an $\bd{Ab}3$-category and let $\left(\{X_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ be a diagram indexed by a small category $I$. Consider a right exact sequence
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=2em,text height=1.5ex, text depth=0.25ex] 
{ \bigoplus_{\alpha \in \Mor(I)}X_{\mathrm{dom}(\alpha)} & & & & & &  \bigoplus_{i\in I}X_i & & & X  \\} ;
\path[->,line width=1.0pt,font=\scriptsize]
(m-1-1) edge node[above] {$\sum_{\alpha \in \Mor(I)}\left(v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)$} (m-1-7)
(m-1-7) edge node[above] {$q $} (m-1-10);
\end{tikzpicture}
\end{center}
where for each $i$ morphism $v_i:X_i \ra \bigoplus_{i\in I}X_i$ is canonical. Define $u_i = q\cdot v_i$ for every $i\in I$. Then $\left(X,\{u_i\}_{i\in I}\right)$ is a colimiting cone of $\left(\{X_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$.
\end{theorem}
\begin{proof}
This is a reformulation of the dual statement to {\cite[page 113, Theorem 1]{Maclane}} and we left it to the reader.
\end{proof}

\begin{definition}
A category $\cC$ is \textit{an $\bd{Ab}4$-category} if it is $\bd{Ab}3$-category and small direct sums in $\cC$ are exact.
\end{definition}

\begin{fact}\label{fact:filteredcolimitsexactisab4}
Let $\cC$ be an $\bd{Ab}3$-category and assume that small filtered colimits in $\cC$ are exact. Then $\cC$ is an $\bd{Ab}4$-category.
\end{fact}
\begin{proof}
Finite direct sums are exact in arbitrary abelian categories and a small direct sum is a filtered colimit of finite direct sums taken over finite subsets of the indexing set. Thus the result follows.
\end{proof}

\begin{definition}
Let $\cC$ be an $\bd{Ab}3$-category. If for every object $X$ of $\cC$, every subobject $Y\subseteq X$ and every directed family $\big\{X_i\}_{i\in I}$ of subobjects of $X$ the following formula holds
$$Y\cap \sum_{i\in I}X_i = \sum_{i\in I}Y\cap X_i$$
then we say that $\cC$ is an $\bd{Ab}5$-category.
\end{definition}
\noindent
The next result is very useful.

\begin{theorem}\label{theorem:kernelofastructuremapinab5}
Let $\cC$ be an $\bd{Ab}5$-category and let $\left(\{X_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ be a diagram indexed by a small filtered category $I$. Suppose that $\left(X,\{u_i\}_{i\in I}\right)$ is a colimiting cone of $\left(\{X_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$. Then
$$\ker(u_j) = \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j}\ker(u_{\alpha})$$
for every $j\in I$.
\end{theorem}
\noindent
The following result is a direct consequence of definition of $\bd{Ab}5$-category.

\begin{lemma}\label{lemma:ab5refinement}
Let $\cC$ be an $\bd{Ab}5$-category, $X$ be its object and $\{X_i\}_{i\in I}$ be a filtered family of subobjects of $X$. Assume also that $f:Y\ra X$ is a monomorphism. Then
$$f^{-1}\left(\sum_{i\in I}X_i\right) = \sum_{i\in I}f^{-1}(X_i)$$
\end{lemma}
\begin{proof}[Proof of the lemma]
It suffices to prove that 
$$f(Y)\cap \sum_{i\in I}X_i = \sum_{i\in I}f(Y)\cap X_i$$
From Fact \ref{fact:imageoffilteredinclass} and applying factorization of $I\ra \mathrm{Sub}(X)$ through its image we can view $\{X_i\}_{i\in I}$ as a directed family of subobjects of $X$. Thus the result follows from the fact that $\cC$ is $\bd{Ab}5$.
\end{proof}

\begin{proof}[Proof of the theorem]
Obviously
$$\sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j}\ker(u_{\alpha})\subseteq \ker(u_j)$$
It suffices to prove that the reverse inclusion holds. For every $i$ in $I$ denote by $v_i:X_i\ra \bigoplus_{i\in I}X_i$ the canonical morphism. By Theorem \ref{theorem:descriptionofcolimitsinab3categories} we have 
$$\ker(u_j) = v_j^{-1}\left(\Image\left(\sum_{\alpha \in \Mor(I)}v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right) = v_j^{-1}\left(\Image\left(\sum_{F\subseteq \Mor(I),\,|F|\in \NN}\sum_{\alpha \in F}\left(v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right)\right)$$
Since $\cC$ is $\bd{Ab}5$-category and by Lemma \ref{lemma:ab5refinement}, we deduce that
$$\ker(u_j) = \sum_{F\subseteq \Mor(I),\,|F|\in \NN}v_j^{-1}\left(\Image\left(\sum_{\alpha \in F} v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right)$$
Thus it suffices to verify that 
$$v_j^{-1}\left(\Image\left(\sum_{\alpha \in F} v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right) \subseteq \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j}\ker(u_{\alpha})$$
for every finite subset $F$ of $\Mor(I)$. Fix such $F$ and suppose that $\{i_1,...,i_n\}$ are all objects of $I$, which are either domains or codomains of arrows in $F$. If $j\not \in \{i_1,...,i_n\}$, then 
$$v_j^{-1}\left(\Image\left(\sum_{\alpha \in F} v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right)=0$$
So we may assume that $j\in \{i_1,...,i_n\}$. Consider a finite diagram in $I$ that consists of $\{i_1,...,i_n\}$ and all arrows in $F$. Since $I$ is filtered, there exists a cocone over this diagram. Hence there exist $i_0$ in $I$ and a family of morphisms $\beta_{i_k}:i_k\ra i_0$ for $1\leq k\leq n$ in $I$ such that $\beta_{\mathrm{dom}(\alpha)} = \beta_{\mathrm{cod}(\alpha)}\cdot \alpha$ for every $\alpha \in F$. Define $f_i = 0$ for $i \in I\setminus \{i_1,...,i_n\}$ and $f_i = u_{\beta_i}$ for $i\in \{i_1,...,i_n\}$. Let $f = \sum_{i\in I}f_i$. We have a commutative square
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=4em,text height=1.5ex, text depth=0.25ex] 
{ X_j                   &   X_{i_0}              \\
  \bigoplus_{i\in I}X_i &  \bigoplus_{i\in I}X_i \\};
\path[->,line width=1.0pt,font=\scriptsize] 
(m-2-1) edge node[below] {$f $} (m-2-2)
(m-1-1) edge node[above] {$u_{\beta_j} $} (m-1-2)
(m-1-1) edge node[left] {$v_j $} (m-2-1)
(m-1-2) edge node[right] {$v_{i_0} $} (m-2-2);
\end{tikzpicture}
\end{center}
and hence it follows that
$$u_{\beta_j}\left(v_j^{-1}\left(\Image\left(\sum_{\alpha \in F} v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right)\right)\subseteq  v_{i_0}^{-1}\left(f\left(\Image\left(\sum_{\alpha \in F} v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right)\right)=$$
$$= v_{i_0}^{-1}\left(\Image\left(\sum_{\alpha \in F}v_{i_0}\cdot \left(u_{\beta_{\mathrm{cod}(\alpha)}}\cdot u_{\alpha}- u_{\beta_{\mathrm{dom}(\alpha)}}\right)\right)\right)=0$$
and thus
$$v_j^{-1}\left(\Image\left(\sum_{\alpha \in F} v_{\mathrm{cod}(\alpha)}\cdot u_{\alpha}-v_{\mathrm{dom}(\alpha)}\right)\right)\subseteq \ker(u_{\beta_j})$$
This finishes the proof.
\end{proof}

\begin{theorem}\label{theorem:ab5isthesameasexactnessoffilteredcolimits}
Let $\cC$ be an $\bd{Ab}3$-category. Then the following are equivalent.
\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=1.5em]
\item $\cC$ is an $\bd{Ab}5$-category.
\item Small filtered colimits are exact in $\cC$.
\end{enumerate}
\end{theorem}
\begin{proof}
The nontrivial part is $\textbf{(i)}\Rightarrow \textbf{(ii)}$. Let $I$ be a small filtered category and 
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{\bigg\{0 &X_i' &  X_i   &X_i''  &          0\bigg\}_{i\in I}         \\} ;
\path[->,font=\scriptsize]  
(m-1-1) edge node[auto] {$ $} (m-1-2)
(m-1-2) edge node[auto] {$r_i $} (m-1-3)
(m-1-3) edge node[auto] {$p_i $} (m-1-4)
(m-1-4) edge node[auto] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
be a diagram of exact sequences indexed by $I$. We denote by $\left(\{X_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ and $\left(\{X'_i\}_{i\in I},\{v_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ appropriate slices of this $I$-indexed diagram. Consider a complex
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{0&X'&  X  &X''&          0           \\} ;
\path[->,font=\scriptsize]  
(m-1-1) edge node[auto] {$ $} (m-1-2)
(m-1-2) edge node[auto] {$r $} (m-1-3)
(m-1-3) edge node[auto] {$p $} (m-1-4)
(m-1-4) edge node[auto] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
where $X'=\mathrm{colim}_{i\in I}X'_i$, $X=\mathrm{colim}_{i\in I}X_i$, $X''=\mathrm{colim}_{i\in I}X''_i$, $r=\mathrm{colim}_{i\in I}r_i$ and $p=\mathrm{colim}_{i\in I}p_i$. Clearly the complex is right exact. It suffices to prove that $r$ is a monomorphism. For $i\in I$ denote by $v_i:X'_i\ra X$, $u_i:X_i\ra X$ structural morphisms. Fix $j\in I$ and consider $Z = v_j^{-1}(\ker(r))$. Then $r_j(Z)\subseteq \ker(u_j)$ and hence by Theorem \ref{theorem:kernelofastructuremapinab5} we derive that
$$r_j(Z) \subseteq \sum_{\alpha \in \Mor(I),\, \mathrm{dom}(\alpha)=j}\ker(u_{\alpha})$$
Note that for every $\alpha \in \Mor(I)$ with $\mathrm{dom}(\alpha) = j$ we have $r_j^{-1}\left(\ker(u_{\alpha})\right)\subseteq \ker(v_{\alpha})$. Indeed, this follows easily from the fact that $r_{\mathrm{cod}(\alpha)}$ is an monomorphism. Thus by Lemma \ref{lemma:ab5refinement}, Theorem \ref{theorem:kernelofastructuremapinab5} and the fact that preimages preserve intersections we deduce that
$$Z = r_{j}^{-1}\left(r_j(Z)\right) =r_j^{-1}\left(\sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j}r_j(Z)\cap \ker(u_{\alpha})\right)=\sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j}Z\cap r_j^{-1}\left(\ker(u_{\alpha})\right)\subseteq $$
$$\subseteq \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j} Z\cap \ker(v_{\alpha}) = Z\cap \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha) = j} \ker(v_{\alpha}) = Z\cap \ker(v_j)$$
and this implies that $Z\subseteq \ker(v_j)$. Since $Z = v_j^{-1}(\ker(r))$, we deduce that 
$$0 = v_j(Z) = \ker(r)\cap \ker(v_j)$$
Now $\{\Image(v_i)\}_{i\in I}$ is a filtered complete family of subobjects of $X'$. Hence by Lemma \ref{lemma:ab5refinement} we deduce that
$$\ker(r) = \ker(r)\cap \sum_{i\in I}\Image(v_i) = \sum_{i\in I}\ker(r)\cap \Image(v_i) = 0$$
and thus $r$ is a monomorphism.
\end{proof}

\begin{corollary}
Every $\bd{Ab}5$-category is $\bd{Ab}4$-category.
\end{corollary}
\begin{proof}
This is a consequence of Fact \ref{fact:filteredcolimitsexactisab4} and Theorem \ref{theorem:ab5isthesameasexactnessoffilteredcolimits}.
\end{proof}

\section{Finite type and finite presentation objects in $\bd{Ab}5$-categories}
\noindent
In this section we investigate properties of finite type objects and related notion of finitely presented objects in $\bd{Ab}5$-categories.

\begin{proposition}\label{proposition:finitetypeandextensions}
Let $\cC$ be an $\bd{Ab}5$-category and consider a short exact sequence
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{  0&  X''  & X & X' &0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge  node[above] {$ $} (m-1-2)
(m-1-2) edge  node[above] {$ $} (m-1-3)
(m-1-3) edge  node[above] {$f$} (m-1-4)
(m-1-4) edge  node[above] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
Then the following assertions hold.
\begin{enumerate}[label=\emph{\textbf{(\arabic*)}}, leftmargin=1.5em]
\item If $X$ is of finite type, then $X'$ is of finite type.
\item If $X''$ and $X'$ are of finite type, then $X$ is of finite type.
\end{enumerate} 
\end{proposition}
\begin{proof}
For the proof of \textbf{(1)} consider a complete filtered family $\{X'_i\}_{i\in I}$ of subobjects of $X'$. Then $\{f^{-1}(X'_i)\}_{i\in I}$ is a complete filtered family of subobjects of $X$. Now $X$ is of finite type. Thus there exists $i_0\in I$ such that $f^{-1}(X'_{i_0})=X$. Hence $X' = X'_{i_0}$. This shows that $X'$ is of finite type.\\
Now we prove \textbf{(2)}. Let $\{X_i\}_{i\in I}$ be a complete filtered family of subobjects of $X$. Since $\cC$ is an $\bd{Ab}5$-category, we derive that $\{X''\cap X_i\}_{i \in I}$ is a complete filtered family of subobjects of $X''$. Moreover, $\{f(X_i)\}_{i\in I}$ is a complete filtered family of subobjects of $X'$. Since both $X'$ and $X''$ are of finite type, there exists $i_0$ and $i_1$ in $I$ such that $X''=X''\cap X_{i_0}$ and $X' =f(X_{i_1})$. Suppose that there are morphisms $i_0\ra i_2$ and $i_1\ra i_2$ for some $i_2\in I$. Then $X'' = X''\cap X_{i_2}$ and $X' = f(X_{i_2})$. This implies that $X = X_{i_2}$ and hence $X$ is of finite type.
\end{proof}

\begin{proposition}\label{proposition:filteredcolimitsandfinitetype}
Let $\cC$ be an $\bd{Ab}5$-category and $X$ be a finite type object of $\cC$. Then for every diagram $\left(\{X'_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ indexed by a small filtered category the canonical morphism
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ \mathrm{colim}_{i\in I}\Mor_{\cC}\left(X,X'_i\right) &  \Mor_{\cC}\left(X,\mathrm{colim}_{i\in I}X'_i\right) \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2);
\end{tikzpicture}
\end{center}
is a monomorphism of abelian groups.
\end{proposition}
\begin{proof}
Denote $\mathrm{colim}_{i\in I}X'_i$ by $X'$ and by $u_i:X'_i\ra X'$ the canonical morphism for every $i\in I$. Fix morphisms $g:X\ra X'_{i}$ for some $i\in I$. Assume that $u_{i}\cdot g = 0$. Then $g(X)\subseteq \ker(u_{i})$ and in addition $g(X)$ is of finite type as the image of $X$ (Proposition \ref{proposition:finitetypeandextensions}). We have
$$\ker(u_{i}) = \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha)=i}\ker(u_{\alpha})$$
by Theorem \ref{theorem:kernelofastructuremapinab5}. We use the fact that $\cC$ is $\bd{Ab}5$ to derive that
$$g(X)= g(X)\cap \ker(u_{i}) = \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha)=i}g(X)\cap \ker(u_{\alpha})$$
Since $g(X)$ is of finite type and $\{g(X)\cap \ker(u_{\alpha})\}_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha)=i}$ is a complete filtered family of subobjects of finite type object $g(X)$, we derive that there exists morphism $\alpha:i\ra i_0$ such that 
$$g(X) = g(X)\cap \ker(u_{\alpha})\subseteq \ker(u_{\alpha})$$
Then $u_{\alpha}\cdot g = 0$ and his implies that $g$ represents zero class in $\mathrm{colim}_{i\in I}\Mor_{\cC}\left(X,X'_i\right)$.
\end{proof}

\begin{definition}
Let $\cC$ be an abelian category and $X$ be an object of $\cC$. Suppose that the following two conditions hold.
\begin{enumerate}[label=\textbf{(\arabic*)}, leftmargin=1.5em]
\item $X$ is of finite type.
\item If $f:X'\ra X$ is an epimorphism in $\cC$ and $X'$ is of finite type, then $\ker(f)$ is of finite type.
\end{enumerate}
Then we say that $X$ is \textit{of finite presentation}.
\end{definition}

\begin{proposition}\label{proposition:finitelypresentedinab5}
Let $\cC$ be an $\bd{Ab}5$-category and $X$ be an object of $\cC$. Then the following assertions are equivalent.
\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=1.5em]
\item $X$ is of finite presentation.
\item There exists an object of finite presentation $X'$ in $\cC$ and an epimorphism $f:X'\ra X$ with kernel of finite type.
\end{enumerate} 
\end{proposition}
\begin{proof}
For the proof of $\textbf{(i)}\Rightarrow\textbf{(ii)}$ it suffices to consider an epimorphism $1_X:X\ra X$ with trivial kernel.\\
Assume that \textbf{(ii)} holds. Fix an epimorphism $f:X'\ra X$ such that $X'$ is of finite presentation and $\ker(f)$ is of finite type. By Proposition \ref{proposition:finitetypeandextensions} object $X$ is of finite type. Next suppose that $g:X''\ra X$ is an epimorphism with $X''$ of finite type. Consider a commutative diagram
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{{}       &\ker(f')   &  \ker(f)   \\
\ker(g')  &Z          &  X'       \\
\ker(g)   &X''        &  X        \\} ;
\path[->,line width=1.0pt,font=\scriptsize] 
(m-1-2) edge node[above] {$\cong$} (m-1-3)
(m-2-1) edge node[left] {$\cong$} (m-3-1)
(m-2-2) edge node[above] {$g'$} (m-2-3)
(m-3-2) edge node[below] {$g$} (m-3-3)
(m-2-2) edge node[left] {$f'$} (m-3-2)
(m-2-3) edge node[right] {$f$} (m-3-3);
\path[right hook->,line width=1.0pt,font=\scriptsize]
(m-1-2) edge node[left] {$ $} (m-2-2)
(m-1-3) edge node[right] {$ $} (m-2-3) 
(m-2-1) edge node[left] {$ $} (m-2-2)
(m-3-1) edge node[left] {$ $} (m-3-2);
\end{tikzpicture}
\end{center}
in which bottom right square is cartesian. The fact that canonical morphisms between kernels in the diagram are isomorphisms follows because $\cC$ is an abelian category. Note also that every row and column of the diagram is a short exact sequence in $\cC$. Since $\ker(f)$ is of finite type, we derive that $\ker(f')$ is of finite type and hence by Proposition \ref{proposition:finitetypeandextensions} object $Z$ is of finite type as an extension of finite type objects $\ker(f')$ and $X''$. Next according to the fact that $X'$ is of finite presentation and $g':Z\ra X'$ is an epimorphism with $Z$ of finite type, we deduce that $\ker(g')$ is of finite type. But $\ker(g')\cong \ker(g)$ and hence $\ker(g)$ is of finite type. This implies that $X$ is of finite presentation. Thus $\textbf{(ii)}\Rightarrow\textbf{(i)}$.
\end{proof}

\begin{theorem}\label{theorem:finitelypresentedandfilteredcolimits}
Let $\cC$ be an $\bd{Ab}5$-category and $X$ be an object of $\cC$. Consider the following statements.
\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=1.5em]
\item $X$ is of finite presentation.
\item For every diagram $\left(\{X'_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ indexed by a small filtered category the canonical morphism
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ \mathrm{colim}_{i\in I}\Mor_{\cC}\left(X,X'_i\right) &  \Mor_{\cC}\left(X,\mathrm{colim}_{i\in I}X'_i\right) \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2);
\end{tikzpicture}
\end{center}
is an isomorphism of abelian groups.
\end{enumerate} 
\end{theorem}
\begin{proof}
We prove that $\textbf{(i)}\Rightarrow \textbf{(ii)}$. Denote $\mathrm{colim}_{i\in I}X'_i$ by $X'$ and suppose that $u_i:X'_i\ra X'$ is the canonical morphism for every $i\in I$. Fix a morphism $f:X\ra X'$ in $\cC$. Let $\left(\{X_i\}_{i\in I},\{v_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ be a diagram obtained by pulling back diagram $\left(\{X'_i\}_{i\in I},\{u_{\alpha}\}_{\alpha \in \Mor(I)}\right)$ along $f$. This means that for every $i\in I$ we have a cartesian square
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{     X_i     &  X'_i        \\
     X       &  X'  \\} ;
\path[->,line width=1.0pt,font=\scriptsize] 
(m-1-1) edge node[above] {$f_i $} (m-1-2)
(m-2-1) edge node[below] {$f $} (m-2-2)
(m-1-1) edge node[left] {$v_i $} (m-2-1)
(m-1-2) edge node[right] {$u_i$} (m-2-2);
\end{tikzpicture}
\end{center}
and if $\alpha:i\ra j$ is a morphism in $I$, then $f_j\cdot v_{\alpha} = u_{\alpha}\cdot f_i$. In $\bd{Ab}5$-category filtered colimits commute with pullbacks (Theorem \ref{theorem:ab5isthesameasexactnessoffilteredcolimits}). Hence $X = \mathrm{colim}_{i\in I}X_i$.  In particular, we have $X = \sum_{i\in I}v_i(X_i)$ and $\{v_i(X_i)\}_{i\in I}$ is a filtered family of subobjects of $X$. Next as $X$ is of finite type, we deduce that there exists $i_0\in I$ such that $X = v_{i_0}(X_{i_0})$. Let $\{Z_k\}_{k\in K}$ be a complete filtered family of subobjects of $X_{i_0}$ that consists of objects of finite type ($\cC$ is locally finite). Then $\{v_{i_0}(Z_k)\}_{k\in K}$ is a complete filtered family of subobjects of $X$. Thus there exists $k_0\in K$ such that $v_{i_0}(Z_{k_0})=X$. This implies that there exists finite type subobject $X''$ of $X_{i_0}$ such that $X= v_{i_0}(X'')$. Since $X$ is of finite presentation, we derive that $X''\cap \ker(v_{i_0})$ is of finite type. On the other hand
$$\ker(v_{i_0}) = \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha)=i_0}\ker(v_{\alpha})$$
by Theorem \ref{theorem:kernelofastructuremapinab5}. Again we use the fact that $\cC$ is $\bd{Ab}5$ to derive that
$$X''\cap \ker(v_{i_0}) = \sum_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha)=i_0}X''\cap \ker(v_{\alpha})$$
Since $X''\cap \ker(v_{i_0})$ is of finite type and $\{X''\cap \ker(v_{\alpha})\}_{\alpha \in \Mor(I),\,\mathrm{dom}(\alpha)=i_0}$ is a complete filtered family of subobjects of $X''\cap \ker(v_{i_0})$, we derive that there exists morphism $\alpha:i_0\ra i_1$ such that 
$$X''\cap \ker(v_{i_0}) = X''\cap \ker(v_{\alpha})$$
Let $X''' = v_{\alpha}(X'')$ be a subobject of $X_{i_1}$. Since $v_{i_0} = v_{i_1}\cdot v_{\alpha}$, we derive that $v_{i_1}$ induces an isomorphism $X'''\cong X$. Therefore, $v_{i_1}:X_{i_1}\ra X$ admits a section $s:X\ra X_{i_1}$. Let $g = f_{i_1}\cdot s$. Then $g:X\ra X'_{i_1}$ is a morphism such that $u_{i_1}\cdot g = f$ and this implies that the canonical morphism
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ \mathrm{colim}_{i\in I}\Mor_{\cC}\left(X,X'_i\right) &  \Mor_{\cC}\left(X,X'\right) \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2);
\end{tikzpicture}
\end{center}
is surjective. The injectivity follows from Proposition \ref{proposition:filteredcolimitsandfinitetype}.\\
Now let us prove that $\textbf{(ii)}\Rightarrow \textbf{(i)}$. Since $\cC$ is locally finite, there exists a complete filtered family $\{X_i\}_{i\in I}$ of subobjects of $X$ and these family consists of objects of finite type in $\cC$. Applying \textbf{(ii)} to this particular filtered diagram we deduce that there exists $i_0$ in $I$ and a morphism $g:X\ra X_{i_0}$ such that $g$ composed with $X_{i_0}\hookrightarrow X$ is $1_X$. This implies that $g$ is an isomorphism and $X = X_{i_0}$. Hence $X$ is of finite type. Thus in order to prove that $X$ is of finite presentation it suffices to check that an epimorphism $f:X'\ra X$ with $X'$ of finite type has the kernel of finite type. For this let $\{K_i\}_{i\in I}$ be a complete filtered family of subobjects of $\ker(f)$ that consists of objects of finite type in $\cC$. We define $X_i = X'/K_i$ for every $i\in I$. Then $\{X_i\}_{i\in I}$ gives rise to a canonical $I$-indexed diagram such that we have an identification
$$X'/\ker(f) = \mathrm{colim}_{i\in I}X_i$$
We apply \textbf{(ii)} to this $I$-indexed diagram and obtain $i_0$ in $I$ together with a morphism $s:X\ra X_{i_0}$ such that $s$ composed with the canonical epimorphism $X_{i_0}\ra X'/\ker(f)$ yields and isomorphism $X \cong X'/\ker(f)$ induced by $f$. This implies that $X_{i_0} \cong X \oplus \left(\ker(f)/K_{i_0}\right)$. Since both $X_{i_0}$ and $X$ are of finite type, we deduce that $\ker(f)/K_{i_0}$ is of finite type. Moreover, $K_{i_0}$ is of finite type. This proves that $\ker(f)$ is of finite type.
\end{proof}

\section{Applications to modules over a ring}
\noindent
Let $R$ be a ring. We denote by $\Mod(R)$ the category of left $R$-modules.

\begin{fact}\label{fact:sumsaredirectedunionsforsubmodules}
Let $M$ be a left $R$-module and $\{M_i\}_{i\in I}$ be a filtered family of submodules of $M$. Then
$$\sum_{i\in I}M_i = \bigcup_{i\in I}M_i$$
\end{fact}
\begin{proof}
Left to the reader.
\end{proof}

\begin{proposition}\label{proposition:modulesarelocallyfinite}
$\Mod(R)$ is a locally finite category and an object $M$ of $\Mod(R)$ is of finite type if and only if $M$ is finitely generated left $R$-module.
\end{proposition}
\begin{proof}
Suppose that $M$ is a finitely generated left $R$-module and $\{M_i\}_{i\in I}$ is a complete filtered family of submodules of $M$. By Fact \ref{fact:sumsaredirectedunionsforsubmodules} we derive that $$M = \bigcup_{i\in I}M_i$$
Let $m_1$,...,$m_n$ be generators of $M$. Then for every $1\leq j\leq n$ there exists $i_j\in I$ such that $m_j\in M_{i_j}$. According to the fact that $I$ is filtered there exists $i_0\in I$ and morphisms $i_j\ra i_0$ for every $1\leq j\leq n$. This implies that $m_j\in M_{i_0}$ for every $1\leq j\leq n$ and hence $M = M_{i_0}$. Therefore, $M$ is an object of finite type in $\Mod(R)$.\\ 
Let $M$ be an arbitrary left $R$-module and $\cF$ be a family of its finitely generated submodules. Since sum of two finitely generated submodules of a given module is finitely generated, we deduce that $\cF$ is directed. Moreover, $M$ together with embeddings $N\hookrightarrow M$ for $N\in \cF$ is the colimit of $\cF$. Therefore, every object in $\Mod(R)$ admits a complete filtered family of subobjects of finite type. Now suppose that $M$ itself is of finite type in $\Mod(R)$. Then $M = N$ for some $N\in \cF$ and hence $M$ is finitely generated.
\end{proof}

\begin{proposition}\label{proposition:modulesformab5category}
$\Mod(R)$ is an $\bd{Ab}5$-category.
\end{proposition}
\begin{proof}
Fix an $R$-module $M$. Let $\{M_i\}_{i\in I}$ be a directed family of its submodules and $N\subseteq M$ be a submodule. By Fact \ref{fact:sumsaredirectedunionsforsubmodules} we have
$$N\cap \sum_{i\in I}M_i = N\cap \bigcup_{i\in I}M_i = \bigcup_{i\in I}\left(N\cap M_i\right) = \sum_{i\in I}N\cap M_i$$ 
\end{proof}

\begin{corollary}\label{corollary:characterizationoffinitelypresentedmodules}
Let $M$ be a left $R$-module. Then the following statements are equivalent.
\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=1.5em]
\item There exist $n$, $m\in \NN$ and a right exact sequence
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=2em,text height=1.5ex, text depth=0.25ex] 
{   R^{\oplus m}  & R^{\oplus n} & M &0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge  node[above] {$ $} (m-1-2)
(m-1-2) edge  node[above] {$ $} (m-1-3)
(m-1-3) edge  node[above] {$ $} (m-1-4);
\end{tikzpicture}
\end{center}
\item $M$ is a finitely presented object of $\Mod(R)$.
\end{enumerate}
\end{corollary}
\begin{proof}
Note that $\Hom_R(R^{\oplus n},-):\Mod(R)\ra \Ab$ is naturally isomorphic with the functor $|-|^{\oplus n}:\Mod(R)\ra \Ab$ that sends each left $R$-module to a direct sum of $n$-copies of its underlying abelian group. By Propositions \ref{proposition:modulesarelocallyfinite} and \ref{proposition:modulesformab5category}, Theorem \ref{theorem:finitelypresentedandfilteredcolimits} and the fact that $|-|^{\oplus n}$ preserves all colimits, we derive that left $R$-module $R^{\oplus n}$ is finitely presented in $\Mod(R)$. Now according to Propositions \ref{proposition:modulesformab5category} and \ref{proposition:finitelypresentedinab5} we deduce that $\textbf{(i)}\Leftrightarrow \textbf{(ii)}$. The converse $\textbf{(ii)}\Rightarrow \textbf{(i)}$ holds by Proposition \ref{proposition:modulesarelocallyfinite} and definition of finitely presented object in abelian category. 
\end{proof}

\begin{example}[Finitely generated module that is not finitely presented]
Let $A$ be a commutative ring and $R = A[x_0,x_1,...]$ be a polynomial $A$-algebra with infinitely many free variables. Denote by $\ideal{a}$ the ideal $\sum_{i\in \NN}R\cdot x_i$ of $R$. Then $M = R/\ideal{a}$ is a finitely generated (even cyclic) $R$-module. On the other hand the kernel of the canonical epimorphism $R\ra R/\ideal{a}=M$ is $\ideal{a}$ and hence it is not a finitely generated $R$-module. Thus $M$ is not finitely presented but finitely generated.
\end{example}

\section{Noetherian and Artinian objects}

\begin{definition}
Let $\cC$ be an abelian category. An object $X$ of $\cC$ is called \textit{noetherian} (\textit{artinian}) if for every ascending (descending) chain $\{X_n\}_{n\in \NN}$ of its subobjects, there exists $n_0\in \NN$ such that $X_n = X_{n_0}$ for $n\geq n_0$.
\end{definition}
\noindent
First we note the following elementary result.

\begin{fact}\label{fact:noetherian_is_dual_to_artinian}
Let $\cC$ be an abelian category. Then noetherian objects in $\cC^{\mathrm{op}}$ are artinian in $\cC$ and vice versa.
\end{fact}
\noindent
The next result is useful for certain applications.

\begin{proposition}\label{proposition:characterization_of_noetherian_objects}
Let $\cC$ be an abelian category and let $X$ be a noetherian (artinian) object of $\cC$. Then for every nonempty class $\cF$ of subobjects of $X$ there exists a maximal (minimal) element contained in this family.
\end{proposition}
\begin{proof}
We present the proof for noetherian case (the proof for artinian case can be obtained by passing to the opposite category). Let $\cF$ be a class of subobjects of a noetherian object $X$. We construct an ascending chain $\{X_n\}_{n\in \NN}$ of subobjects of $\cF$ as follows. We pick $X_0\in \cF$. Suppose that $X_0\subseteq ...\subseteq X_n$ are defined. If $X_n$ is a maximal element of $\cF$, then we set $X_{n+1} = X_n$. Otherwise we pick $X_{n+1}\in \cF$ such that $X_n\subsetneq X_{n+1}$. Since $X$ is noetherian, there exists $n_0\in \NN$ such that $X_n = X_{n_0}$ for $n\geq n_0$. This implies that $X_{n_0}$ is maximal element in $\cF$.
\end{proof}

\begin{definition}
Let $\cC$ be an abelian category and let $\cS$ be its full subcategory. Suppose that for every exact sequence in $\cC$
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ 0 & X' & X & X'' & 0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$ $} (m-1-3)
(m-1-3) edge node[above] {$ $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
we have $X\in \cS$ if and only if $X',X''\in \cS$. Then $\cS$ is called \textit{a thick subcategory of $\cC$}.
\end{definition}

\begin{proposition}\label{proposition:noetherian_and_artinian_is_thick}
Let $\cC$ be an abelian category. Then subcategory of $\cC$ consisting of noetherian (artinian) objects is thick.
\end{proposition}
\begin{proof}
We present the proof for noetherian case (the proof for artinian case can be obtained by passing to the opposite category). Consider a short exact sequence
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ 0 & X' & X & X'' & 0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$i $} (m-1-3)
(m-1-3) edge node[above] {$p $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
Suppose first that $X$ is noetherian. Consider an ascending chain $\{X'_n\}_{n\in \NN}$ of subobjects of $X'$. Then it is also an ascending chain of subobjects of $X$. Hence there exists $n_0\in \NN$ such that $X'_n = X'_{n_0}$ for $n\geq n_0$. This implies that $X'$ is noetherian. Now suppose that $\{X''_n\}_{n\in \NN}$ is an ascending chain of subobjects of $X''$. Then $\{p^{-1}(X''_n)\}_{n\in \NN}$ is an ascending chain of subobjects of $X$. Hence there exists $n_0\in \NN$ such that $p^{-1}(X''_n) = p^{-1}(X''_{n_0})$ for $n\geq n_0$. This shows that $X''_n = X''_{n_0}$ for $n\geq n_0$. Thus $X''$ is noetherian.\\
Next suppose that both $X'$ and $X''$ are noetherian. Let $\{X_n\}_{n\in \NN}$ be an ascending chain of subobjects of $X$. Then $\{i^{-1}(X_n)\}_{n\in \NN}$ and $\{p(X_n)\}_{n\in \NN}$ are ascending chains of subobjects of $X'$ and $X''$, respectively. Since $X'$ and $X''$ are noetherian, we derive that there exists $n_0\in \NN$ such that $i^{-1}(X_n) = i^{-1}(X_{n_0})$ and $p(X_n) = p(X_{n_0})$ for $n\geq n_0$. This implies that $X_n = X_{n_0}$ for $n\geq n_0$ and hence $X$ is noetherian.
\end{proof}

\section{Objects of finite length}
It is interesting to investigate the case when object is both noetherian and artinian. We need the following definition first.

\begin{definition}
Let $\cC$ be an abelian category. An nonzero object $X$ of $\cC$ is \textit{irreducible} if the only subobjects of $X$ are zero and $X$ itself.
\end{definition}

\begin{proposition}\label{proposition:characterization_of_objects_of_flength}
Let $\cC$ be an abelian category and let $X$ be an object of $\cC$. Then the following assertions are equivalent.
\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=1.5em]
\item $X$ is both noetherian and artinian.
\item There exists a chain
$$0=X_0\subseteq X_1\subseteq X_2\subseteq ...\subseteq X_n = X$$
of subobjects of $X$ such that for each $0\leq i\leq n-1$ the quotient $X_{i+1}/X_i$ is an irreducible object of $\cC$.
\end{enumerate}
\end{proposition}
\noindent
For the proof we will need the following.

\begin{lemma}\label{lemma:minimal_subobject_above_given_subobject}
Let $\cC$ be an abelian category and let $X$ be an artinian object of $\cC$. Suppose that $X'\subsetneq X$ is a subobject. Then there exists a minimal subobject $X''$ of $X$ such that $X'\subsetneq X''$.
\end{lemma}
\begin{proof}[Proof of the lemma]
It suffices to apply Proposition \ref{proposition:characterization_of_noetherian_objects} to a nonempty class
$$\cF = \big\{Z\,\big|\,Z\mbox{ is a subobject of }X\mbox{ strictly containing }X'\big\}$$
\end{proof}

\begin{proof}[Proof of the proposition]
We prove $\textbf{(i)}\Rightarrow \textbf{(ii)}$. Suppose that $X$ is both noetherian and artinian. Since $X$ is artinian, by Lemma \ref{lemma:minimal_subobject_above_given_subobject} we construct a chain
$$0 = X_0\subsetneq X_1\subsetneq ...$$
such that for each $i$ the quotient $X_{i+1}/X_i$ is irreducible object of $\cC$. Indeed, starting from $X_0 = 0$ we apply the following procedure. If $X_i$ was constructed and $X_i\subsetneq X$, then by Lemma \ref{lemma:minimal_subobject_above_given_subobject} we pick $X_{i+1}$ as a minimal subobject of $X$ such that $X_i\subsetneq X_{i+1}$. Note that the construction of this ascending chain of subobjects of $X$ cannot go indefinitely, because $X$ is noetherian. Hence we deduce that
$$0 = X_0\subsetneq X_1\subsetneq ...\subsetneq X_n = X$$
for some $n\in \NN$.\\
Now we prove that $\textbf{(ii)}\Rightarrow \textbf{(i)}$. Assume that there exists a chain
$$0=X_0\subseteq X_1\subseteq X_2\subseteq ...\subseteq X_n = X$$
of subobjects of $X$ such that for each $0\leq i\leq n-1$ the quotient $X_{i+1}/X_i$ is an irreducible object of $\cC$. Then we prove by finite induction that each $X_i$ is both noetherian and artinian. This clearly holds for $X_0$. Suppose that it holds for $X_i$ for some $i<n$. Then there exists a canonical short exact chain
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ 0 & X_i & X_{i+1} & X_{i+1}/X_i & 0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$ $} (m-1-3)
(m-1-3) edge node[above] {$ $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
Since $X_i$ is both noetherian and artinian and moreover, $X_{i+1}/X_i$ as an irreducible object is both noetherian and artinian, we derive by Proposition \ref{proposition:noetherian_and_artinian_is_thick} that $X_{i+1}$ is both noetherian and artinian.
\end{proof}

\begin{definition}
Let $\cC$ be an abelian category and let $X$ be an object of $\cC$. A chain
$$0=X_0 \subseteq X_1\subseteq ...\subseteq X_n = X$$
of subobjects such that $X_{i+1}/X_i$ is an irreducible object for $0 \leq i \leq n-1$ is called \textit{a composition series of $X$}. Irreducible objects
$$X_1/X_0,X_2/X_1,...,X_n/X_{n-1}$$
are called \textit{factors of the composition series}.
\end{definition}
\noindent
Let $X$ be an object of a category $\cC$. Then we denote by $[X]_{\cong}$ the isomorphism class of $X$.

\begin{theorem}[Jordan-Hölder]\label{theorem:JordanHölder_equivalence}
Let $\cC$ be an abelian category and let $X$ be its object. Suppose that
$$0=X_0 \subseteq X_1\subseteq ...\subseteq X_n = X$$
and
$$0=X'_0 \subseteq X'_1\subseteq ...\subseteq X'_m = X$$
are composition series of $X$. Then multisets (sets which allow repetitions)
$$\{[X_1/X_0]_{\cong},[X_2/X_1]_{\cong},...,[X_n/X_{n-1}]_{\cong}\},\,\{[X'_1/X'_0]_{\cong},[X'_2/X'_1]_{\cong},...,[X'_m/X'_{m-1}]_{\cong}\}$$
are equal.
\end{theorem}
\begin{proof}
The proof goes by induction on $n$. If $n=1$, then $X$ is irreducible and the result holds.\\
Suppose now $n\geq 1$. Consider object $Z = X/X_1$. Then $Z_i = X_{i+1}/X_1$ for $0\leq i\leq n-1$. Clearly
$$0=Z_0\subseteq Z_1\subseteq ...\subseteq Z_{n-1} = Z$$
is a composition series of $Z$ and we have
$$\{[Z_1/Z_0]_{\cong},[Z_2/Z_1]_{\cong},...,[Z_{n-1}/Z_{n-2}]_{\cong}\} = \{[X_2/X_1]_{\cong},[X_3/X_2]_{\cong},...,[X_n/X_{n-1}]_{\cong}\}$$
Consider now objects $W_i = X'_i/(X'_i\cap X_1)$ for $0\leq i\leq m$. Then we have a chain of subobjects
$$0 = W_0 \subseteq W_1\subseteq ...\subseteq W_m = Z$$
of $Z$. Pick $i = 0,1,...,m-1$. We have a commutative diagram
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{0 & X'_i & X'_{i+1}  & X'_{i+1}/X'_i & 0 \\
 0 & W_i  & W_{i+1}   & W_{i+1}/W_i   & 0  \\};
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$ $} (m-1-3)
(m-1-3) edge node[above] {$ $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5)

(m-2-1) edge node[above] {$ $} (m-2-2)
(m-2-2) edge node[above] {$ $} (m-2-3)
(m-2-3) edge node[above] {$ $} (m-2-4)
(m-2-4) edge node[above] {$ $} (m-2-5)

(m-1-2) edge node[left] {$p_i $} (m-2-2)
(m-1-3) edge node[left] {$p_{i+1} $} (m-2-3)
(m-1-4) edge node[right] {$q_{i+1} $} (m-2-4);
\end{tikzpicture}
\end{center}
in $\cC$ with exact rows and canonical epimorphism as vertical arrows. By Snake Lemma ({\cite[page 11]{weibel1995introduction}}) we derive that there exists an exact sequence
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=1em,text height=1.5ex, text depth=0.25ex] 
{ 0 & \Ker(p_i) & \Ker(p_{i+1}) & \Ker(q_{i+1}) & \Coker(p_i) & \Coker(p_{i+1}) & \Coker(q_{i+1}) & 0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$ $} (m-1-3)
(m-1-3) edge node[above] {$ $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5)
(m-1-5) edge node[above] {$ $} (m-1-6)
(m-1-6) edge node[above] {$ $} (m-1-7)
(m-1-7) edge node[above] {$ $} (m-1-8);
\end{tikzpicture}
\end{center}
Since $\Ker(p_i) = X'_i\cap X_1,\,\Ker(p_{i+1}) = X'_{i+1}\cap X_1$ and $p_i,p_{i+1},q_{i+1}$ are epimorphism, we deduce that this exact sequence induces a short exact sequence
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ 0 & X'_i\cap X_1 & X'_{i+1}\cap X_1  & \Ker(q_{i+1}) & 0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$ $} (m-1-3)
(m-1-3) edge node[above] {$ $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
Pick now $k$ such that $X'_i\cap X_1 = 0$ for $i\leq k$ and $X'_i\cap X_1 = X_1$ for $k<i$. Thus for $i \neq k$ we have $\Ker(q_{i+1}) = 0$ and hence $X'_{i+1}/X'_i$ is isomorphic to $W_{i+1}/W_i$. Moreover, for $i = k$ we have $X_1\cong \Ker(q_{k+1})$. Hence $X_1 \cong X'_{k+1}/X'_k$ and $W_{k+1}/W_k = 0$. We define
$$Z'_i = \begin{cases}W_i & \mbox{ for }i\leq k\\
W_{i+1} & \mbox{ for }k<i
\end{cases}
$$
for $0\leq i\leq m-1$. Then
$$0 = Z'_0 \subseteq Z'_1\subseteq ...\subseteq Z'_{m-1}=Z$$
is a composition series and
$$\{[Z'_1/Z'_0]_{\cong},[Z'_2/Z'_1]_{\cong}....,[Z'_m/Z'_{m-1}]_{\cong}\} = \{[X'_2/X'_1]_{\cong},[X'_3/X'_2]_{\cong},...,[X'_m/X'_{m-1}]_{\cong}\}$$
Now by induction we have
$$\{[Z'_1/Z'_0]_{\cong},[Z'_2/Z'_1]_{\cong}....,[Z'_m/Z'_{m-1}]_{\cong}\} = \{[Z_1/Z_0]_{\cong},[Z_2/Z_1]_{\cong}....,[Z_m/Z_{m-1}]_{\cong}\}$$
Since $X'_{k+1}/X'_k \cong X_1$, we deduce that
$$\{[X'_1/X'_0]_{\cong},[X'_2/X'_1]_{\cong},...,[X'_m/X'_{m-1}]_{\cong}\} = \{[Z'_1/Z'_0]_{\cong},[Z'_2/Z'_1]_{\cong}....,[Z'_m/Z'_{m-1}]_{\cong}\}\cup \{[X'_{k+1}/X'_k]_{\cong}\} =$$
$$=\{[Z_1/Z_0]_{\cong},[Z_2/Z_1]_{\cong}....,[Z_m/Z_{m-1}]_{\cong}\}\cup  \{[X_1]_{\cong}\} = \{[X_1/X_0]_{\cong},[X_2/X_1]_{\cong},...,[X_n/X_{n-1}]_{\cong}\}$$
\end{proof}


\begin{definition}
Let $\cC$ be an abelian category and let $X$ be an object of $\cC$ that has a composition series (or equivalently that $X$ is both noetherian and artinian). Then we say that $X$ is \textit{an object of finite length}.
\end{definition}

\begin{corollary}\label{corollary:irreducible_class_multiplicity}
Let $X$ be an object of finite length in an abelian category $\cC$ and let $\lambda$ be an isomorphism type of some irreducible object in $\cC$. Suppose that
$$0=X_0 \subseteq X_1\subseteq ...\subseteq X_n = X$$
is a composition series. Let $l_{\lambda}(X)$ be the number of elements of the multiset
$$\{[X_1/X_0]_{\cong},[X_2/X_1]_{\cong},...,[X_n/X_{n-1}]_{\cong}\}$$
equal to $\lambda$. Then $l_{\lambda}(X)$ does not depend on the composition series of $X$.
\end{corollary}
\begin{proof}
This is a consequence of the fact that multisets of isomorphism types of factors are equal for any two composition series of $X$ (Theorem \ref{theorem:JordanHölder_equivalence}).
\end{proof}

\begin{definition}
Let $X$ be an object of finite length in an abelian category $\cC$. Suppose that $\Lambda$ is a class of isomorphism types of irreducible objects in $\cC$. For a given $\lambda \in \Lambda$ we define $l_{\lambda}(X)$ as in Corollary \ref{corollary:irreducible_class_multiplicity} above. We call it \textit{the multiplicity of $\lambda$ in $X$}. Next we define
$$l(X) = \sum_{\lambda\in \Lambda}l_{\lambda}(X)\in \NN$$
(note that the components of the sum on the right hand side are almost all equal to zero) and we call this number \textit{the length of $X$}.
\end{definition}

\begin{proposition}\label{proposition:length_and_multipicity_as_morphism_from_Grothendieck_group}
Suppose that
\begin{center}
\begin{tikzpicture}
[description/.style={fill=white,inner sep=2pt}]
\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex] 
{ 0 & X' & X & X'' & 0 \\} ;
\path[->,line width=1.0pt,font=\scriptsize]  
(m-1-1) edge node[above] {$ $} (m-1-2)
(m-1-2) edge node[above] {$i $} (m-1-3)
(m-1-3) edge node[above] {$p $} (m-1-4)
(m-1-4) edge node[above] {$ $} (m-1-5);
\end{tikzpicture}
\end{center}
is a short exact sequence of finite length objects in an abelian category $\cC$. Then for every isomorphism type $\lambda$ of some irreducible object $X$ in $\cC$ we have
$$l_{\lambda}(X) = l_{\lambda}(X')+l_{\lambda}(X'')$$
\end{proposition}
\begin{proof}
Since $X'$ and $X''$ are of finite length, we fix composition series
$$0=X'_0 \subseteq X'_1\subseteq ...\subseteq X'_m = X',\,0= X''_0 \subseteq X''_1\subseteq ...\subseteq X''_k = X''$$
Then
$$0 = i(X'_0)\subseteq ...\subseteq i(X'_m)\subseteq p^{-1}(X''_1)\subseteq ...\subseteq p^{-1}(X''_k) = X$$
is a composition series for $X$. By Corollar \ref{corollary:irreducible_class_multiplicity} we can calculate $l_{\lambda}(X)$ by means of this compositon series. This implies that
$$l_{\lambda}(X) = l_{\lambda}(X')+l_{\lambda}(X'')$$
\end{proof}

\section{Semisimple objects}

\begin{proposition}\label{proposition:finite_type_objects_admit_maximal_proper_subobjects}
Let $\cC$ be a well-powered $\bd{Ab}3$-category and let $X$ be a nonzero object of $\cC$ of finite type. Then there exists a subobject $X'$ of $X$ such that $X/X'$ is irreducible.
\end{proposition}
\begin{proof}
Let $\cF$ be the class of proper subobjects of $X$. That is $\cF = \mathrm{Sub}(X)\setminus \{X\}$. This class is nonempty, because $X$ is nonzero. Since $\cC$ is well-powered, we derive that $\cF$ is a set. Consider a linearly ordered set $I$ and a set $\{X_i\}_{i\in I}$ of elements of $\cF$ such that $X_{i_1}\subseteq X_{i_2}$ for $i_1\leq i_2$ in $I$. If $X = \sum_{i\in I}X_i$, then there exists $i_0\in I$ such that $X = X_{i_0}$. This is a contradiction with the fact that $X_{i_0}$ in an element of $\cF$. This implies that $\sum_{i\in I}X_i$ is an element of $\cF$. Hence chains of elements of $\cF$ admit upper bounds. By Zorn's lemma we deduce that there exists a maximal element $X'$ in $\cF$. Then $X'$ is a proper and maximal subobject of $X$. This is equivalent with $X/X'$ being irreducible.
\end{proof}

\begin{definition}
Let $\cC$ be an abelian category and let $X$ be an object of $\cC$. If every subobject $X'$ of $X$ is a direct summand of $X$, then we say that $X$ is \textit{completely reducible}.
\end{definition}

\begin{proposition}\label{proposition:completely_reducible_class_closed_under_subobjects}
Let $\cC$ be an abelian category. Then the class of completely reducible objects of $\cC$ is closed under subobjects.
\end{proposition}
\begin{proof}
Let $X$ be a completely reducible object of $\cC$. Consider its subobject $Y$. Assume that $Y'$ is a subobject of $Y$. Since $X$ is completely reducible, we derive that there exists a subobject $X'$ such that $X = Y' + X'$ and $X'\cap Y'=0$. Since subobject lattices in abelian categories are modular (it is a consequence of Theorem \ref{theorem:Freyd_Mitchel_embedding}), we derive that
$$Y = Y\cap X = Y\cap (Y' +  X') = Y' + Y\cap X'$$
Moreover, $Y'\cap (Y\cap X')=0$. Thus $Y'$ is a direct summand of $Y$. This proves that $Y$ is completely reducible.
\end{proof}

\begin{definition}
Let $\cC$ be an abelian category and let $X$ be an object of $\cC$. If $X$ is the sum of its irreducible subobjects, then we say that $X$ is \textit{semisimple}.
\end{definition}

\begin{proposition}\label{proposition:semisimple_closed_under_quotients}
Let $\cC$ be an abelian category. Then the class of semisimple objects of $\cC$ is closed under quotient objects.
\end{proposition}
\begin{proof}
Let $X$ be a semisimple object of $\cC$. Consider an epimorphism $q:X\ra X'$. Since $q$ preserves sums of subobjects and an epimorphic image of irreducible subobject is either zero or irreducible, we derive that $X'$ is semisimple.
\end{proof}

\begin{theorem}
Let $\cC$ be an abelian category and let $X$ be an object of $\cC$. Consider the following statements.
\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=1.5em]
\item $X$ is semisimple.
\item $X$ is completely reducible.
\end{enumerate}
If $\cC$ is a well-powered $\bd{Ab}5$-category, then $\emph{\textbf{(i)}}\Rightarrow \emph{\textbf{(ii)}}$. If in addition $X$ is locally finite, then $\emph{\textbf{(ii)}} \Rightarrow \emph{\textbf{(i)}}$.
\end{theorem}
\begin{proof}
Assume that $\cC$ is a well-powered $\bd{Ab}5$-category. We prove $\textbf{(i)}\Rightarrow \textbf{(ii)}$. Consider a subobject $X'$ of $X$. Consider a class
$$\cF=\big\{Y\,\big|\,Y\mbox{ is a subobject of }X\mbox{ and }Y\cap X' =0\big\}$$
Since $\cC$ is well-powered, the class $\cF$ is a set. Since $\cC$ is an $\bd{Ab}5$-category, Zorn's lemma implies that the set $\cF$ ordered by inclusion of subobjects has a maximal element $X''$. Now consider an irreducible subobject $K$ of $X$. If $K\cap (X'+X'')= 0$, then $Y = X'' + K$ is a subobject of $X$ containing $X''$ and satisfying $Y\cap X' = 0$. In particular, $Y$ is an element of $\cF$. By maximality of $X''$ in $\cF$, we derive that $Y = X''$ and hence $K$ is a contained in $X''$. This is a contradiction because then $K= K \cap (X' + X'') = 0$ and irreducible objects are nonzero. Thus for every irreducible subobject $K$ of $X$ we have $K\cap (X'+X'')\neq 0$. This implies that $K = K\cap (X'+X'')$ and hence $K$ is contained in $X'+X''$. Since $X$ is a sum of its irreducible subobjects, we derive that $X = X' + X''$. Therefore, $X'$ is a direct summand of $X$.\\
Suppose that in addition $X$ is locally finite. Now we prove that $\textbf{(ii)}\Rightarrow \textbf{(i)}$. Suppose that $X'$ is a sum of all irreducible subobjects of $X$. Since $X'$ is a direct summand of $X$, we derive that there exists a subobject $X''$ of $X$ such that $X'\cap X''=0$ and $X = X'+X''$. Assume that $X''$ is nonzero. Since $X$ is locally finite and $X''$ is isomorphic with $X/X'$, we deduce that $X''$ is locally finite. Consider a nonzero subobject $Y$ of $X''$ of finite type. By Proposition \ref{proposition:finite_type_objects_admit_maximal_proper_subobjects} we deduce that there exists a subobject $Y'$ of $Y$ such that $Y/Y'$ is irreducible. According to Proposition \ref{proposition:completely_reducible_class_closed_under_subobjects} we deduce that $Y$ is completely reducible as a subobject of $X$. Hence there exists a subobject $Y''$ of $Y$ such that $Y = Y' + Y''$ and $Y'\cap Y''=0$. Clearly $Y''\cong Y/Y'$ and hence $Y''$ is irreducible subobject of $Y$. Since $Y$ is contained in $X''$, we deduce that $X''$ admits an irreducible subobject. This is contradiction with $X'\cap X'' = 0$. This proves that $X''$ must be zero and hence $X$ is a sum of its irreducible subobjects.
\end{proof}


















\small
\bibliographystyle{apalike}
\bibliography{../zzz}

\end{document}