\input ../pree.tex

\begin{document}

\title{Daniell-Kolmogorov extension theorem}
\date{}
\maketitle

\section{Introduction}
\noindent
In these notes we prove Daniell-Kolmogorov extension theorem, which concerns existence of probability measures on infinite products. Daniell-Kolmogorov theorem is a fundamental result of advanced probability theory. It implies the existence of various stochastic processes.

\section{Daniell-Kolmogorov extension theorem}

\begin{theorem}[Daniell-Kolmogorov]\label{theorem:Daniell_Kolmogorov_extension}
Let $T$ be a set and let $(X_t,\cF_t)_{t\in T}$ be a set of measurable spaces. For every $t\in T$ we fix a Hausdorff topology $\tau_t$ on $X_t$. Suppose that the following assertions hold.
\begin{enumerate}[label=\emph{\textbf{(\arabic*)}}, leftmargin=*]
\item For every finite subset $F$ of $T$ there exists a probability measure $\mu_F$ on the measurable space $(X_F,\cF_F)$. The space is defined by
$$X_F = \prod_{t\in F}X_t,\,\cF_F = \bigotimes_{t\in F}\cF_t$$
\item If $F_1\subseteq F_2$ are finite subsets of $T$ and $\pi_{F_2,F_1}:X_{F_2}\ra X_{F_1}$ is the projection, then $\left(\pi_{F_2,F_1}\right)_*\mu_{F_2} = \mu_{F_1}$.
\item The measure $\mu_F$ is inner regular with respect to the product topology $\tau_F = \prod_{t\in \cF}\tau_t$.
\end{enumerate}
Then there exists a unique probability measure $\mu$ on the space $(X,\cF)$ where 
$$X = \prod_{t\in T}X_t,\,\cF = \bigotimes_{t\in T}\cF_t$$
such that for every finite subset $F$ of $T$ we have
$$\mu_F = \left(\pi_F\right)_*\mu$$
for the projection $\pi_F:X\ra X_F$.
\end{theorem}
\noindent
For the proof we need some notation. First recall that $\cF = \bigotimes_{t\in T}\cF_t$ is the $\sigma$-algebra generated by subsets of $X$ of the form $\pi_F^{-1}(B)$ where $B\in \cF_F$. We call such sets \textit{cylinders}. We prove now some lemma which relies on inner regularity assumption.

\begin{lemma}\label{lemma:inner_regularity_in_Daniell_Kolmogorov_extension}
Let $\{F_n\}_{n\in \NN}$ be a nondecreasing sequence of finite subsets of $T$ and let $\{D_n\}_{n\in \NN}$ be sets such that $D_n\in \cF_{F_n}$ and $\{\pi_{F_n}^{-1}(D_n)\}_{n\in \NN}$ form a nonincreasing family of cylinders in $X$. Assume that there exists $\epsilon > 0$ such that 
$$\mu_{F_n}(D_n) \geq \epsilon$$
for every $n\in \NN$. Then there exists a sequence $\{K_n\}_{n\in \NN}$ such that the following assertions are satisfied.
\begin{enumerate}[label=\emph{\textbf{(\arabic*)}}, leftmargin=*]
\item $K_n$ is a subset of $D_n$ for every $n\in \NN$.
\item $K_n$ is compact with respect to $\tau_{F_n}$ for every $n\in \NN$.
\item $\{\pi^{-1}_{F_n}(K_n)\}_{n\in \NN}$ is a nonincreasing sequence.
\item $\mu_{F_n}(K_n) \geq \frac{\epsilon}{2}$ for every $n\in \NN$.
\end{enumerate}
\end{lemma}
\begin{proof}[Proof of the lemma]
We prove slightly stronger statement. More precisely, we prove that there exists a sequence $\{K_n\}_{n\in \NN}$ satisfying all the assertions in the statement and the additional assertion that
$$\mu_{F_n}(D_n)\leq \mu_{F_n}(K_n) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right)$$
for all $n\in \NN$. The proof goes by induction. For $D_0$ we pick its subset $K_0$ compact with respect to $\tau_{F_0}$ such that 
$$\mu_{F_0}(D_0) \leq \mu_{F_0}(K_0) + \frac{\epsilon}{4}$$
Suppose that for some $n\in \NN$ there is a sequence $\{K_m\}_{m\leq n}$ such that $K_m\subseteq D_m$, $K_m$ is compact with respect to $\tau_{F_{m}}$, the sequence $\{\pi^{-1}_{F_m}(K_m)\}_{m\leq n}$ is nonincreasing and
$$\mu_{F_m}(D_m)\leq \mu_{F_m}(K_m) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^m}\right)$$
for every $m\leq n$. Write $\pi_{F_{n+1},F_n}^{-1}(K_n) = \tilde{K}_n$ and $\pi_{F_{n+1},F_n}^{-1}(D_n) = \tilde{D}_n$. Then
$$\mu_{F_{n+1}}(\tilde{D}_n) = \mu_{F_n}(D_n) \leq \mu_{F_n}(K_n) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right) = \mu_{F_{n+1}}(\tilde{K}_n) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right)$$
and $\tilde{K}_n \subseteq \tilde{D}_n$. Moreover, by assumption $D_{n+1}\subseteq \tilde{D}_n$. Thus we have
$$\mu_{F_{n+1}}\left(D_{n+1} \setminus (\tilde{K}_{n}\cap D_{n+1})\right) = \mu_{F_{n+1}}(D_{n+1} \setminus \tilde{K}_{n}) \leq \mu_{F_{n+1}}(\tilde{D}_{n} \setminus \tilde{K}_{n}) \leq \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right)$$
and hence
$$\mu_{F_{n+1}}\left(D_{n+1}\right) - \mu_{F_{n+1}}\left(\tilde{K}_{n}\cap D_{n+1}\right) \leq \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right)$$
Now we pick a subset $K_{n+1}$ of $\tilde{K}_{n}\cap D_{n+1}$ such that $K_{n+1}$ is compact with respect to $\tau_{F_{n+1}}$ and 
$$\mu_{F_{n+1}}(\tilde{K}_{n}\cap D_{n+1}) \leq \mu_{F_{n+1}}(K_{n+1}) + \frac{\epsilon}{4}\cdot \frac{1}{2^{n+1}}$$
Then
$$\mu_{F_{n+1}}\left(D_{n+1}\right) \leq \mu_{F_{n+1}}\left(\tilde{K}_{n}\cap D_{n+1}\right) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right) \leq $$
$$\leq \mu_{F_{n+1}}(K_{n+1}) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^n}\right) + \frac{\epsilon}{4}\cdot \frac{1}{2^{n+1}} = \mu_{F_{n+1}}(K_{n+1}) + \frac{\epsilon}{4}\cdot \left(1 + \frac{1}{2} + ... + \frac{1}{2^{n+1}}\right)$$
Hence we can extend sequence $\{K_m\}_{m\leq n}$ by adding $K_{n+1}$ and our inductive proof is completed.
\end{proof}
\noindent
The next lemma is purely topological.

\begin{lemma}\label{lemma:topological_part_kolmogorov_extension}
Let $\cS$ be a countable set. Suppose that $\{F_n\}_{n\in \NN}$ is a nondecreasing sequence of finite subsets of $\cS$ such that $\cS = \bigcup_{n\in \NN}F_n$ and assume that $\{X_s\}_{s\in \cS}$ is a family of Hausdorff topological spaces. If $\{K_n\}_{n\in \NN}$ is a sequence of nonempty sets such that $K_n$ is a compact subset of $\prod_{s\in F_n}X_s$ for every $n\in \NN$ and 
$$\bigg\{K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg\}_{n\in \NN}$$
is a nondecreasing sequence of subsets of $\prod_{s\in \cS}X_s$, then
$$\bigcap_{n\in \NN}\bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg) \neq \emptyset$$
\end{lemma}
\begin{proof}[Proof of the lemma]
For each $s\in \cS$ we choose $n_s\in \NN$ such that $s \in F_{n_s}$. We define 
$$K_{s} = \pi_{s}\bigg(K_{n_s}\times \prod_{s\in \cS\setminus F_{n_s}}X_s\bigg)$$
where $\pi_s:\prod_{s\in \cS}X_s\ra X_s$ is the projection. Then by assumptions $K_{s}$ is a compact and nonempty subset of $X_s$. We define $K$ as a product of $\prod_{s\in \cS}K_s$. By Tychonoff's theorem $K$ is compact and nonempty subset of $\prod_{s\in \cS}X_s$. Fix now $n\in \NN$. Define $m = \max\left(n,\max_{s\in F_n}n_s\right)$. Then
$$\pi_s\bigg(K_{m}\times \prod_{s\in \cS\setminus F_{m}}X_s\bigg) \subseteq \pi_s\bigg(K_{n_s}\times \prod_{s\in \cS\setminus F_{n_s}}X_s\bigg)\subseteq K_s$$
for every $s\in F_n$. Thus
$$K_{m}\times \prod_{s\in \cS\setminus F_{m}}X_s \subseteq \prod_{s\in F_n}K_s\times \prod_{s\in \cS\setminus F_n}X_s$$
Since $m\geq n$, we have
$$K_{m}\times \prod_{s\in \cS\setminus F_{m}}X_s \subseteq \bigg(\prod_{s\in F_n}K_s\times \prod_{s\in \cS\setminus F_n}X_s\bigg)\cap \bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg)$$
Together with $K_m\neq \emptyset$ this implies that
$$\bigg(\prod_{s\in F_n}K_s\times \prod_{s\in \cS\setminus F_n}X_s\bigg)\cap \bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg)\neq \emptyset$$
We infer that
$$K\cap \bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg) \neq \emptyset$$
Therefore, the sequence 
$$\bigg\{K\cap \bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg)\bigg\}_{n\in \NN}$$
is nondecreasing and consists of closed nonempty subsets of a compact topological space $K$. Hence it has nonempty intersection. This completes the proof that
$$\bigcap_{n\in \NN}\bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg) \neq \emptyset$$
\end{proof}

\begin{proof}[Proof of the theorem]
Fix a finite subset $F$ of $T$ and for $B$ in $\cF_F$ we set
$$\mu\left(\pi^{-1}_F(B)\right) = \mu_F(B)$$
Note that this makes $\mu$ into a function defined on the family of all cylinders. Moreover, it is clear that the family of all cylinders is an algebra of subsets of $X$ and $\mu$ defined in such a way is an additive function. We claim that $\mu$ is also $\sigma$-additive. For this pick a nondecreasing sequence $\{F_n\}_{n\in \NN}$ of finite subsets of $T$ and a sequence $\{B_n\}_{n\in \NN}$ of sets such that $B_n\in \cF_{F_n}$ for every $n\in \NN$ and the sequence $\{\pi_{F_n}^{-1}(B_n)\}_{n\in \NN}$ is nondecreasing. Moreover, we assume that there exists a finite subset $F$ of $T$ and a set $B \in \cF_{F}$ such that 
$$\pi_{F}^{-1}(B) = \bigcup_{n\in \NN}\pi_{F_n}^{-1}(B_n)$$
We define $D_n = \pi_{F\cup F_n,F}^{-1}(B)\setminus \pi_{F\cup F_n,F_n}^{-1}(B_n)$ for every $n\in \NN$. Then $D_n \in \cF_{F\cup F_n}$ satisfies
$$\pi^{-1}_{F\cup F_n}(D_n) = \pi_F^{-1}(B)\setminus \pi^{-1}_{F_n}(B_n)$$
Hence the sequence $\{\pi^{-1}_{F\cup F_n}(D_n)\}_{n\in \NN}$ is nonincreasing and 
$$\emptyset = \bigcap_{n\in \NN}\pi^{-1}_{F\cup F_n}(D_n)$$
Now suppose that there exists $\epsilon > 0$ such that $\mu\left(\pi^{-1}_{F\cup F_n}(D_n)\right) \geq \epsilon$ for every $n\in \NN$. Then by Lemma \ref{lemma:inner_regularity_in_Daniell_Kolmogorov_extension} there exists a sequence $\{K_n\}_{n\in \NN}$ such that the following assertions are satisfied.
\begin{enumerate}[label=\textbf{(\arabic*)}, leftmargin=*]
\item $K_n$ is a subset of $D_n$ for every $n\in \NN$.
\item $K_n$ is compact with respect to $\tau_{F\cup F_n}$ for every $n\in \NN$.
\item $\{\pi^{-1}_{F\cup F_n}(K_n)\}_{n\in \NN}$ is a nonincreasing sequence.
\item $\mu_{F\cup F_n}(K_n) \geq \frac{\epsilon}{2}$ for every $n\in \NN$.
\end{enumerate}
In particular, assertion \textbf{(4)} implies that $K_n$ is nonempty for every $n\in \NN$. We set $\cS = \bigcup_{n\in \NN}\left(F\cup F_n\right)$ and apply Lemma \ref{lemma:topological_part_kolmogorov_extension} to $\cS$, $\{(X_s,\tau_s)\}_{s\in \cS}$ and $\{K_n\}_{n\in \NN}$. We obtain
$$\bigcap_{n\in \NN}\bigg(K_n\times \prod_{s\in \cS\setminus F_n}X_s\bigg) \neq \emptyset$$
which implies that
$$\bigcap_{n\in \NN}\pi^{-1}_{F\cup F_n}(K_n) \neq \emptyset$$
This is contradiction, since
$$\bigcap_{n\in \NN}\pi^{-1}_{F\cup F_n}(K_n) \subseteq \bigcap_{n\in \NN}\pi^{-1}_{F\cup F_n}(D_n) = \emptyset$$
Therefore, we proved that 
$$\lim_{n\ra +\infty}\mu(\pi^{-1}_{F\cup F_n}(D_n)) = 0$$
and hence
$$\mu(\pi_F^{-1}(B)) = \lim_{n\ra +\infty}\mu(\pi^{-1}_{F_n}(B_n))$$
This proves the claim that $\mu$ is $\sigma$-additive. According to {\cite[Theorem 3.3]{Introduction_to_measure_theory}} additive function $\mu$ defined on the algebra of cylinders can be extended to a unique probability measure on $\cF$. Since no confusion can arise, we denote this probability measure by $\mu$. Then $\mu:\cF\ra [0,1]$ is a unique probability measure such that 
$$\mu\left(\pi^{-1}_F(B)\right) = \mu_F(B)$$
for every finite subset $F$ of $T$ and $B\in \cF_F$.
\end{proof}


\small
\bibliographystyle{apalike}
\bibliography{../zzz}
\end{document}