% arara: indent: {overwrite: yes, silent: yes}
\documentclass[10pt]{amsart}
\input ../pree.tex

\begin{document}
\title{Topological groups}
\date{}
\maketitle

\section{Introduction}
\noindent
In these notes we study topological groups which are very important objects in analysis. We depend \cite{Topological_Spaces}, \cite{Filters_in_topology} and \cite{Uniform_Spaces}.

In the first section we introduce topological groups and study their topological properties. Next section studies left and right uniform structures on topological groups and their relation to topological notions. We also introduce two-sided uniform structures and study their properties. The last section is devoted to completions of topological groups.

\section{Topological groups}

\begin{definition}
	A group together with a topology such that group operations are continuous is a \textit{topological group}.
\end{definition}

\begin{definition}
	Let $G,H$ be a topological groups. A map $f:G\ra H$ which is continuous homomorphism is a \textit{morphism} of topological groups.
\end{definition}

\begin{definition}
	Let $i:G\hookrightarrow H$ be a morphism of topological groups and a topological embedding. Then $i$ is an \textit{embedding} of topological groups.
\end{definition}
\noindent
Let $G$ be a topological group. For a subset $S$ of $G$ we define
$$S^{-1} = \big\{x^{-1}\,\big|\,x\in S\,\big\}$$

\begin{definition}
	A subset $S$ of a topological group $G$ is \textit{symmetric} if $S$ and $S^{-1}$ coincide.
\end{definition}

\begin{fact}\label{fact:symmetric_neighborhoods_of_unit_generate_base}
	Let $G$ be a topological group and let $O$ be an open neighborhood of identity in $G$. Then there exists an open and symmetric neighborhood $Q$ of identity in $G$ such that $Q\subseteq O$.
\end{fact}
\begin{proof}
	Since $(-)^{-1}:G\ra G$ is homeomorphism, $O \cap O^{-1}$ is symmetric and open neighborhood of identity contained in $O$.
\end{proof}

\begin{fact}\label{fact:closures_in_topological_groups}
	Let $G$ be a topological group and let $\cO$ be the family of all open and symmetric neighborhoods of identity in $G$. Then the following assertions hold.
	\begin{enumerate}[label=\emph{\textbf{(\arabic*)}}, leftmargin=*]
		\item Let $A$ be the subset of $G$. Then
		      $$\bd{cl}(A) = \bigcap_{O \in \cO}OA = \bigcap_{O \in \cO}OAO = \bigcap_{O \in \cO}AO$$
		\item If $H$ is a subgroup of $G$, then $\bd{cl}(H)$ is a subgroup of $G$.
		\item If $N$ is a normal subgroup of $G$, then $\bd{cl}(N)$ is a normal subgroup of $G$.
	\end{enumerate}
\end{fact}
\begin{proof}
	Left for the reader as an exercise.
\end{proof}

\begin{theorem}\label{theorem:quotients_of_topological_groups}
	Let $G$ be a topological group and let $N$ be its normal subgroup. Consider the quotient map $q:G\twoheadrightarrow G/N$ in the category of groups and equip $G/N$ with quotient topology. Then the following assertions holds.
	\begin{enumerate}[label=\emph{\textbf{(\arabic*)}}, leftmargin=*]
		\item $q$ is an open map.
		\item $G/N$ is a topological group and $q$ is a homomorphism of groups.
		\item Let $f:G\ra H$ be a continuous homomorphism of topological groups and suppose that $N \subseteq \Ker(f)$. Then there exists a unique continuous homomorphism $g:G/N\ra H$ such that $g\cdot q = f$.
		\item $N$ is a closed in $G$ if and ony if $G/N$ is Hausdorff.
	\end{enumerate}
\end{theorem}
\noindent
For the proof we need the following result.

\begin{lemma}\label{lemma:Hausdorff_topological_groups}
	Let $G$ be a topological group. Then $G$ is Hausdorff if and only if identity subgroup of $G$ is closed.
\end{lemma}
\begin{proof}[Proof of the lemma]
	If $G$ is Hausdorff, then each singleton subset of $G$ is closed. Hence identity subgroup of $G$ is closed.

	Conversely, assume that identity subgroup in $G$ is closed. Pick two distinct elements $g_1,g_2 \in G$. Since $G$ is a topological group, the map
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=4em, column sep=4em,text height=1.5ex, text depth=0.25ex]
			{ G\times G     & G\times G & G \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ 1_G\times (-)^{-1} $} (m-1-2)
			(m-1-2) edge node[above] {$ \cdot_G $} (m-1-3);
		\end{tikzpicture}
	\end{center}
	is continuous. Hence there exists an open neighborhood $O$ of identity in $G$ such that $g_1g_2^{-1}\not \in OO^{-1}$. Then
	$$Og_1\cap Og_2 = \emptyset$$
	Thus $G$ is a Hausdorff topological space.
\end{proof}

\begin{proof}[Proof of the theorem]
	Fix an open subset $Q$ of $G$, then the set
	$$q^{-1}\left(q\left(Q\right)\right) = QN$$
	is open. According to the fact that $q:G \twoheadrightarrow G/N$ is a quotient topological map, we infer that $q(Q)$ is open in $G/N$. Hence $q$ is an open map and the proof of \textbf{(1)} is completed.

	Since $q$ is open, we derive that $q\times q$ is open. Since squares
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=4em, column sep=5em,text height=1.5ex, text depth=0.25ex]
			{ G \times G     & G   & G   & G                            \\
			  G/N \times G/N & G/N & G/N & G/N \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ \cdot_{G} $} (m-1-2)
			(m-1-1) edge node[left] {$ q\times q $} (m-2-1)
			(m-1-2) edge node[right] {$ q $} (m-2-2)
			(m-2-1) edge node[below] {$ \cdot_{G/N} $} (m-2-2)
			(m-1-3) edge node[above] {$ (-)^{-1}_{G} $} (m-1-4)
			(m-1-3) edge node[left] {$ q $} (m-2-3)
			(m-1-4) edge node[right] {$ q $} (m-2-4)
			(m-2-3) edge node[below] {$ (-)^{-1}_{G/N} $} (m-2-4);
		\end{tikzpicture}
	\end{center}
	are commutative, we deduce that the addition $\cdot_{G/N}:G/N \times G/N \ra G/N$ and the inverse map $(-)^{-1}_{G/N}:G/N\ra G/N$ are continuous. Therefore, $G/N$ is a topological group. It follows that $q$ is a morphism of topological groups and hence \textbf{(2)} holds.

	The assertion \textbf{(3)} describes the universal property which follows easily from \textbf{(2)} and the fact that $q$ is a topological quotient.

	For \textbf{(4)} observe that
	$$N\mbox{ is closed subgroup of }G\,\Leftrightarrow\,\mbox{identity subgroup of }G/N\mbox{ is closed }$$
	Thus it suffices to prove that
	$$\mbox{ identity subgroup of }G/N\mbox{ is closed }\,\Leftrightarrow\,G/N\mbox{ is a Hausdorff topological space}$$
	but this is a consequence of Lemma \ref{lemma:Hausdorff_topological_groups}.
\end{proof}

\section{Uniform structures on topological groups}
\noindent
In this section we introduce uniform structures on topological groups and study their properties.

\begin{fact}\label{fact:left_uniform_structure_on_topological_group}
	Let $G$ be a topological group. For every open and symmetric neighborhood $O$ of identity in $G$ we define
	$$L_O = \big\{(g_1,g_2)\in G\times G\,\big|\,g_1^{-1}g_2 \in O\big\}$$
	The collection of $L_O$ for all open and symmetric neighborhoods of identity in $G$ is a base of the uniform structure on $G$,
\end{fact}
\begin{proof}
	Note that $L_O$ is reflexive and symmetric relation on $G$ for every open and symmetric neighborhood $O$ of identity in $G$.

	Next it is clear that $L_{O_1}\cap L_{O_2} = L_{O_1\cap O_2}$ for open and symmetric neighborhoods $O_1,O_2$ of identity in $G$.

	Pick open and symmetric neighborhood $O$ of identity in $G$. Since the multiplication $G\times G\ra G$ is continuous, there exists an open neighborhood $Q$ of unit in $G$ such that $QQ \subseteq O$. By Fact \ref{fact:symmetric_neighborhoods_of_unit_generate_base} we may assume that $Q$ is symmetric. Hence $L_Q\cdot L_Q \subseteq L_O$.
\end{proof}

\begin{definition}
	Let $G$ be a topological group. Then the uniformity introduced above is \textit{left uniform structure} on $G$.
\end{definition}

\begin{proposition}\label{proposition:left_uniform_structure_on_topological_group}
	Let $G$ be a topological group. Then the following assertions hold.
	\begin{enumerate}[label=\emph{\textbf{(\arabic*)}}, leftmargin=3.0em]
		\item Left uniform structure induces the topology on $G$.
		\item For every $g \in G$ maps
		      $$G\ni x \mapsto gx \in G,\,G\ni x \mapsto xg \in G$$
		      are uniform with respect to left uniform structure on $G$.
	\end{enumerate}
\end{proposition}
\begin{proof}
	Suppose that $O$ is an open and symmetric neighborhood of identity in $G$. Then
	$$L_O(x) = xO^{-1} = xO$$
	for every $x \in G$. Hence $Q \subseteq G$ is open set in topology induced by left uniform structure if and only if
	$$Q = \bigcup_{x\in Q}xO_x$$
	where $O_x$ is an open and symmetric neighborhood of the identity in the original topology of $G$ for each $x \in Q$. Fact \ref{fact:symmetric_neighborhoods_of_unit_generate_base} implies that left uniform structure on induces the original topology on $G$. This completes the proof of \textbf{(1)}.

	For $g \in G$ we denote by $l_g$ and $r_g$ maps $G \ni x \mapsto gx \in G$ and $G \ni x \mapsto xg \in G$, respectively. Then
	$$\left(l_g\times l_g\right)^{-1}(L_O) = L_O,\,\left(r_g \times r_g\right)^{-1}(L_O) = L_{gOg^{-1}}$$
	for every open and symmetric neighborhood $O$ of identity in $G$. Thus $l_g$ and $r_g$ are uniform with respect to left uniform structure on $G$. Hence \textbf{(2)} holds.
\end{proof}

\begin{fact}\label{fact:each_continuous_homomorphism_is_uniform_with_respect_to_left_uniformity}
	Let $f:G\ra H$ be a continuous homomorphism of topological groups and let $O$ be an open and symmetric neighborhood of identity in $G$. Then
	$$\left(f\times f\right)^{-1}(L_O) = L_{f^{-1}(O)}$$
	In particular, $f$ is uniform with respect to left uniform structures on $G$ and $H$.
\end{fact}
\begin{proof}
	We have
	$$\left(f\times f\right)^{-1}(L_O) = \big\{(g_1, g_2) \in G\times G\,\big|\,f(g_1)^{-1}f(g_2) \in O\big\} = L_{f^{-1}(O)}$$
	Thus $f$ is a uniform map.
\end{proof}

\begin{corollary}\label{corollary:topological_groups_embeddings_is_uniform_embedding}
	Let $i:G\hookrightarrow H$ be an embedding of topological groups. Then $i$ is a uniform embedding with respect to left uniformities on $G$ and $H$.
\end{corollary}
\begin{proof}
	Since $i$ is a topological embedding, the map $O\mapsto f^{-1}(O)$ defined which takes open and symmetric neighborhoods of $H$ to open and syummetric neighborhoods of $G$ is surjective. By Fact \ref{fact:each_continuous_homomorphism_is_uniform_with_respect_to_left_uniformity} we derive
	$$L_{i^{-1}(O)} = \left(i\times i\right)^{-1}(L_O)$$
	for every open and symmetric neighborhood of $H$. Hence $i$ induces left uniform structure on $G$. Thus $i$ is an embedding of $G$ into $H$ with respect to left uniform structure.
\end{proof}

\begin{corollary}\label{corollary:uniform_Kolmogorov_quotients_of_topological_groups}
	Let $G$ be a topological group and let $q:G\twoheadrightarrow \tilde{G}$ be the quotient of $G$ with respect to the closure of identity in $G$. If $\tilde{G}$ and $G$ are considered with their left uniform structures, then $q$ is a uniform Kolmogorov quotient.
\end{corollary}
\begin{proof}
	We denote by $1$ the identity of $G$. It follows from Fact \ref{fact:closures_in_topological_groups} that the relation of topological indistinguishability on $G$ is given by cosets of normal subgroup $\bd{cl}\left(\{1\}\right) \subseteq G$. Pick an open and symmetric neighborhood $O$ of identity in $G$. Then $q(O)$ is an open and symmetric neighborhood of identity in $\tilde{G}$ such that $q^{-1}(q(O)) = O$. Thus
	$$\left(q \times q\right)^{-1}(L_{q(O)}) = L_O$$
	Hence we derive that $q$ is a uniform Kolmogorov quotient of $G$.
\end{proof}

\begin{remark}\label{remark:right_uniform_structure}
	Let $G$ be a topological group. For every open and symmetric neighborhood $O$ of identity in $G$ we define
	$$R_O = \big\{(g_1,g_2)\in G\times G\,\big|\,g_1g_2^{-1} \in O\big\}$$
	The collection of $R_O$ for all open and symmetric neighborhoods of identity in $G$ is a base of the \textit{right uniform structure} on $G$.
\end{remark}

\begin{fact}\label{fact:inverse_is_an_isomorphism_of_left_and_right_uniform_structures}
	Let $G$ be a topological group. Then $(-)^{-1}:G\ra G$ is a uniform map between left and right uniformities on $G$.
\end{fact}
\begin{proof}
	Left for the reader as an exercise.
\end{proof}

\begin{definition}
	Let $G$ be a topological group and let $\cF$ be a filter on $G$. If $\cF$ is Cauchy with respect to left (right) uniform structure on $G$, then $\cF$ is \textit{left} (\textit{right}) Cauchy.
\end{definition}

\begin{definition}
	Let $G$ be a topological group which is complete with respect to its left (right) uniform structure. Then $G$ is \textit{left} (\textit{right}) complete.
\end{definition}

\begin{remark}\label{remark:join_of_uniform_structures}
	Let $X$ be a set and let $\fU_1$ and $\fU_2$ be uniform structures on $X$. Then
	$$\fU_1\vee \fU_2 = \big\{U \in \fD_X\,\big|\,U_1\cap U_2\subseteq U\mbox{ for some }U_1\in \fU_1\mbox{ and }U_2\in \fU_2\big\}$$
	is a uniform structure on $X$, which is the least upper bound of $\fU_1$ and $\fU_2$ in the partially ordered set of uniform structures on $X$. Note that if $\fU_1$ and $\fU_2$ induce some topologies on $X$, then $\fU_1\vee \fU_2$ induce the least upper bound of these two topologies in the partial order of all topologies on $X$.
\end{remark}

\begin{definition}
	Let $G$ be a topological group. Then the least upper bound of left and right uniform structures on $G$ is the \textit{two-sided} uniform structure on $G$.
\end{definition}

\begin{fact}\label{fact:inverse_is_an_automorphism_of_two_sided_uniform_structures}
	Let $G$ be a topological group. Then $(-)^{-1}:G\ra G$ is a uniform map with respect to two-sided uniform structure on $G$.
\end{fact}
\begin{proof}
	Left for the reader as an exercise.
\end{proof}

\begin{definition}
	Let $G$ be a topological group and let $\cF$ be a filter on $G$. If $\cF$ is Cauchy with respect to two-sided unifomr structure on $G$, then $\cF$ is \textit{two-sided} Cauchy.
\end{definition}

\begin{fact}\label{fact:two_sided_Cauchy_filters}
	Let $G$ be a topological group and let $\cF$ be a filter on $G$. Then $\cF$ is a two-sided Cauchy filter on $G$ if and only if it is both left and right Cauchy filter on $G$.
\end{fact}
\begin{proof}
	Left for the reader as an exercise.
\end{proof}

\begin{definition}
	Let $G$ be a topological group which is complete with respect to its two-sided uniform structure. Then $G$ is \textit{two-sided} complete.
\end{definition}

\section{Completions of topological groups}
\noindent
In this section we prove two important theorems concerning completions of topological groups. We start by proving two results, which are of independent interest.

\begin{proposition}\label{proposition:Hausdorff_uniform_structure_is_determined_by_dense_subspace}
	Let $Z$ be a uniform space and let $X$ be a set containing $Z$. Then there exists at most one Hausdorff uniform structure on $X$ such that the inclusion $Z\hookrightarrow X$ is uniform embedding with dense image.
\end{proposition}
\begin{proof}
	Let $\fU_1$,$\fU_2$ be Hausdorff uniform structures on $X$ such that $Z\hookrightarrow (X,\fU_1)$ and $Z \hookrightarrow (X.\fU_2)$ are uniform embeddings with dense image. Replacing $\fU_2$ by $\fU_1\vee \fU_2$ we may assume that $\fU_1$ is contained in $\fU_2$. It follows that $1_X$ is the uniform map $(X,\fU_2)\ra (X,\fU_1)$. Let $j_i:(X,\fU_i)\hookrightarrow \overline{X}_i$ be Hausdorff completions of $(X,\fU_i)$ for $i=1,2$. There exists a unique uniform map $\overline{X}_1\dashrightarrow \overline{X}_2$ such that the diagram
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=2.5em, column sep=0.8em,text height=1.5ex, text depth=0.25ex]
			{ \overline{X}_2 &   & \overline{X}_1         \\
			  (X,\fU_2)      &   & (X,\fU_1)                  \\
			                 & Z & \\} ;
			\path[densely dotted,->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$  $} (m-1-3);
			\path[->,line width=0.8pt,font=\scriptsize]

			(m-2-1) edge node[above] {$ 1_X $} (m-2-3);
			\path[right hook->,line width=0.8pt,font=\scriptsize]
			(m-2-1) edge node[left] {$ j_2 $} (m-1-1)
			(m-2-3) edge node[right] {$ j_1 $} (m-1-3)

			(m-3-2) edge node[left] {$  $} (m-2-1)
			(m-3-2) edge node[right] {$  $} (m-2-3);
		\end{tikzpicture}
	\end{center}
	is commutative. Since vertical maps are uniform embeddings with dense images, we derive that both $Z \hookrightarrow \overline{X}_1$ and $Z\hookrightarrow \overline{X}_2$ are Hausdorff completions. Thus the top horizontal arrow is an isomorphism of uniform spaces. It follows that $1_X:(X,\fU_2)\ra (X,\fU_1)$ is an isomorphism of uniform spaces. Hence $\fU_1$ and $\fU_2$ coincide.
\end{proof}

\begin{proposition}\label{proposition:multiplication_preserves_one_sided_Cauchy_filters}
	Let $G$ be a topological group. Then the multiplication $G\times G\ra G$ sends left (right) Cauchy filters on $G\times G$ to left (right) Cauchy filters.
\end{proposition}
\begin{proof}
	We prove proposition for left uniform structures.

	Let $\cF$ be a left Cauchy filter in $G\times G$ and let $O$ be an open and symmetric neighborhood of $G$. Let $\cF_l$ and $\cF_r$ be images of $\cF$ under left and right projections $G\times G\ra G$. Then $\cF_l$ and $\cF_r$ are left Cauchy filters on $G$. Fix open and symmetric neighborhood $Q$ of $G$ such that $QQQ \subseteq O$. There exists $F_r \in \cF_r$ such that
	$$F_r \times F_r \subseteq L_Q$$
	Next fix $x \in F_r$. Next there exists $F_l \in \cF_l$ such that
	$$F_l\times F_l \subseteq L_{xQx^{-1}}$$
	Define $F = F_l\times F_r$ and note that $F \in \cF$. Pick $(l_1,r_1),(l_2,r_2) \in F$. Then
	$$(l_1r_1)^{-1}(l_2r_2) = r_1^{-1}(l_1^{-1}l_2)r_2 = (r_1^{-1}x)x^{-1}(l_1^{-1}l_2)x(x^{-1}r_2) \subseteq Qx^{-1}\left(xQx^{-1}\right)xQ = QQQ \subseteq O$$
	Hence the preimage of $L_O$ under the multiplication of $G$ contains $F \times F$. It follows that the image of $\cF$ under the multiplication of $G$ is a Cauchy filter on $G$.
\end{proof}

\begin{definition}
	Let $G$ be a topological group. A left (right) complete group $\overline{G}$ and an embedding $G \hookrightarrow \overline{G}$ of topological groups with dense image is a \textit{left} (\textit{right}) group completion of $G$.
\end{definition}

\begin{theorem}\label{theorem:criterion_for_existence_of_one_sided_completions}
	Let $G$ be a topological group. Consider the following assertions.
	\begin{enumerate}[label=\emph{\textbf{(\roman*)}}, leftmargin=*]
		\item $G$ admits left completion.
		\item Left Cauchy filters and right Cauchy filters on $G$ coincide.
	\end{enumerate}
	Then $\emph{\textbf{(i)}} \Rightarrow \emph{\textbf{(ii)}}$ and if $G$ is Hausdorff, then $\emph{\textbf{(ii)}} \Rightarrow \emph{\textbf{(i)}}$.
\end{theorem}
\begin{proof}
	Suppose that $G$ admits left completion $i:G \hookrightarrow \overline{G}$. Let $\cF$ be a left Cauchy filter on $G$. We define
	$$\cF^{-1} = \big\{F^{-1}\,\big|\,F\in \cF\big\}$$
	Then there exists a continuous map $\overline{G}\ra \overline{G}$ such that the diagram
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex]
			{  G            & G                                                          \\
			   \overline{G} & \overline{G}                      \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ (-)^{-1} $} (m-1-2)
			(m-2-1) edge node[below] {$  $} (m-2-2);
			\path[right hook->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[left] {$ i $} (m-2-1)
			(m-1-2) edge node[right] {$ i $} (m-2-2);
		\end{tikzpicture}
	\end{center}
	is commutative. From the commutativity it follows that $i(\cF^{-1})$ is convergent in $\overline{G}$. Since $\overline{G}$ is left complete and $i$ is an embedding of uniform spaces by Corollary \ref{corollary:topological_groups_embeddings_is_uniform_embedding}, we derive that $\cF^{-1}$ is a left Cauchy filter on $G$. According to Fact \ref{fact:inverse_is_an_isomorphism_of_left_and_right_uniform_structures} the map $(-)^{-1}$ induces bijection between left and right Cauchy filters on $G$. Thus $\cF$ is a right Cauchy filter on $G$. This completes the proof of $\textbf{(i)}\Rightarrow \textbf{(ii)}$.

	Suppose now that left and right Cauchy filters on $G$ coincide and $G$ is Hausdorff. Let $\overline{G}$ be a Hausdorff and complete uniform space and $i:G\hookrightarrow \overline{G}$ be a uniform embedding with dense image where $G$ is considered with its left uniform structure. Note that $\overline{G}$ is a regular topological space. Since $(-)^{-1}_G:G\ra G$ sends left Cauchy filters to left Cauchy filters by assumption and by the general result on extension of continuous maps described in \cite{Filters_in_topology}, we derive that there exists a unique continuous map $(-)^{-1}_{\overline{G}}$ such that the square
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex]
			{  G            & G                                                          \\
			   \overline{G} & \overline{G}                      \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ (-)^{-1}_G $} (m-1-2)
			(m-2-1) edge node[below] {$ (-)^{-1}_{\overline{G}}  $} (m-2-2);
			\path[right hook->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[left] {$ i $} (m-2-1)
			(m-1-2) edge node[right] {$ i $} (m-2-2);
		\end{tikzpicture}
	\end{center}
	is commutative. Next by Proposition \ref{proposition:multiplication_preserves_one_sided_Cauchy_filters} the multiplication $\cdot_G:G\times G \ra G$ sends left Cauchy filters on $G\times G$ onto left Cauchy filters on $G$. Again the general result on extension of continuous maps in \cite{Filters_in_topology} there exists a unique continuous map $\cdot_{\overline{G}}$ such that the square
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=4em, column sep=4em,text height=1.5ex, text depth=0.25ex]
			{ G\times G                        & G                                     \\
			  \overline{G} \times \overline{G} & \overline{G} \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ \cdot_G $} (m-1-2)
			(m-2-1) edge node[below] {$ \cdot_{\overline{G}} $} (m-2-2);
			\path[right hook->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[left] {$ i\times i $} (m-2-1)
			(m-1-2) edge node[right] {$ i $} (m-2-2);
		\end{tikzpicture}
	\end{center}
	is commutative. Now since $i$ has dense image and $G$ is a topological group and $\overline{G}$ is Hausdorff, we deduce that $\overline{G}$ is a topological group with operations $\cdot_{\overline{G}}$ and $(-)^{-1}_{\overline{G}}$. Moreover, $i$ is an embedding of topological groups. Finally the uniform structure on $\overline{G}$ is the left uniform structure of topological group $\overline{G}$ by Proposition \ref{proposition:Hausdorff_uniform_structure_is_determined_by_dense_subspace}.
\end{proof}


\begin{definition}
	Let $G$ be a topological group. A two-sided complete group $\overline{G}$ and an embedding $G \hookrightarrow \overline{G}$ of topological groups with dense image is a \textit{two-sided} group completion of $G$.
\end{definition}

\begin{theorem}\label{theorem:two_sided_completion_always_exists}
	Let $G$ be a Hausdorff topological group. Then $G$ admits two-sided completion.
\end{theorem}
\begin{proof}
	Let $\overline{G}$ be a Hausdorff and complete uniform space and $i:G\hookrightarrow \overline{G}$ be a uniform embedding with dense image where $G$ is considered with its two-sided uniform structure. Note that $\overline{G}$ is a regular topological space. By Fact \ref{fact:inverse_is_an_automorphism_of_two_sided_uniform_structures} and by general result on extension of uniform maps described in \cite{Uniform_Spaces} we infer that there exists a unique uniform map $(-)^{-1}_{\overline{G}}$ such that the square
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=3em, column sep=3em,text height=1.5ex, text depth=0.25ex]
			{  G            & G                                                          \\
			   \overline{G} & \overline{G}                      \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ (-)^{-1}_G $} (m-1-2)
			(m-2-1) edge node[below] {$ (-)^{-1}_{\overline{G}}  $} (m-2-2);
			\path[right hook->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[left] {$ i $} (m-2-1)
			(m-1-2) edge node[right] {$ i $} (m-2-2);
		\end{tikzpicture}
	\end{center}
	is commutative. Next by Proposition \ref{proposition:multiplication_preserves_one_sided_Cauchy_filters} the multiplication $\cdot_G:G\times G \ra G$ sends left Cauchy filters on $G\times G$ onto left Cauchy filters on $G$. By symmetry it also sends right Cauchy filters to right Cauchy filters. Fact \ref{fact:two_sided_Cauchy_filters} implies that $\cdot_{G}$ sends two-sided Cauchy filters to two-sided Cauchy filters. According to the general result on extension of continuous maps in \cite{Filters_in_topology} there exists a unique continuous map $\cdot_{\overline{G}}$ such that the square
	\begin{center}
		\begin{tikzpicture}
			[description/.style={fill=white,inner sep=2pt}]
			\matrix (m) [matrix of math nodes, row sep=4em, column sep=4em,text height=1.5ex, text depth=0.25ex]
			{ G\times G                        & G                                     \\
			  \overline{G} \times \overline{G} & \overline{G} \\} ;
			\path[->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[above] {$ \cdot_G $} (m-1-2)
			(m-2-1) edge node[below] {$ \cdot_{\overline{G}} $} (m-2-2);
			\path[right hook->,line width=0.8pt,font=\scriptsize]
			(m-1-1) edge node[left] {$ i\times i $} (m-2-1)
			(m-1-2) edge node[right] {$ i $} (m-2-2);
		\end{tikzpicture}
	\end{center}
	is commutative. Now since $i$ has dense image and $G$ is a topological group and $\overline{G}$ is Hausdorff, we deduce that $\overline{G}$ is a topological group with operations $\cdot_{\overline{G}}$ and $(-)^{-1}_{\overline{G}}$. Moreover, $i$ is an embedding of topological groups. Finally the uniform structure on $\overline{G}$ is the two-sided uniform structure of topological group $\overline{G}$ by Proposition \ref{proposition:Hausdorff_uniform_structure_is_determined_by_dense_subspace}.



\end{proof}





















\small
\bibliographystyle{apalike}
\bibliography{../zzz}

\end{document}

